### 2023-05-26 paper/main.tex
Branchful code

\documentclass[11pt]{revtex4}

\usepackage[colorlinks=true]{hyperref}

\begin{document}
\title{Magnetic Phase Transitions in an Aperiodic Monotile}
\author{Jack Dinsmore}
\affiliation{Department of Physics, Stanford University}

\begin{abstract}
  Abstract
\end{abstract}

\maketitle

Hello there.

\end{document}

### 2023-05-26 src/lattice.rs

@@ -1,0 +1,79 @@
+use std::collections::VecDeque;
+
+use crate::spin::{Spin, SmallVec};
+use rand::random;
+
+pub struct Lattice<S: Spin> {
+    data: Vec<S>,
+    neighbors: Vec<SmallVec<usize>>,
+}
+
+impl<S: Spin> Lattice<S> {
+    pub fn new(num_sites: usize, neighbor_func: impl Fn(usize) -> SmallVec<usize>) -> Self {
+        let mut data = Vec::with_capacity(num_sites);
+        let mut neighbors = Vec::with_capacity(num_sites);
+        for site_index in 0..num_sites {
+            data.push(S::start());
+            neighbors.push(neighbor_func(site_index));
+        }
+
+        Self {
+            data,
+            neighbors,
+        }
+    }
+
+    pub fn zero(&mut self) {
+        for item in self.data.iter_mut() {
+            *item = S::start();
+        }
+    }
+
+    pub fn run(&mut self, beta: f64, n_trials: usize) -> f64 {
+        let rolling_avg_count = 500;
+        let runs_per_trial = 16;
+        
+        self.evolve(beta, runs_per_trial * (n_trials - rolling_avg_count));
+        let mut rolling_avg = 0.0;
+        for _ in 0..rolling_avg_count {
+            self.evolve(beta, runs_per_trial);
+            rolling_avg += self.order_param();
+        }
+        rolling_avg / rolling_avg_count as f64
+    }
+
+    fn evolve(&mut self, beta: f64, count: usize) {
+        for _ in 0..count {
+            // 1. Choose random site
+            let start_index = random::<usize>() % self.data.len();
+            let vec = S::random_vec();
+
+            // 2. Mark, and flip.
+            self.data[start_index].flip(&vec);
+            let mut new_sites = VecDeque::new();
+            new_sites.push_back(start_index);
+            let mut marked_sites = vec![false; self.data.len()];
+
+            while let Some(s) = new_sites.pop_front() {
+                // 3. Iterate through neighbors and flip
+                for neighbor in self.neighbors[s].iter_mut() {
+                    if marked_sites[*neighbor] { continue; }
+                    let prob = 1. - (0.0f64.min(2. * beta * self.data[s].dot(&vec) * self.data[*neighbor].dot(&vec))).exp();
+                    if random::<f64>() < prob {
+                        self.data[*neighbor].flip(&vec);
+                        new_sites.push_back(*neighbor);
[…]

### 2023-05-26 src/main.rs

@@ -1,3 +1,41 @@
+#![allow(dead_code)]
+mod lattice;
+mod spin;
+use lattice::Lattice;
+use spin::{Ising, SmallVec};
+
 fn main() {
-    println!("Hello, world!");
+    let row_size = 100;
+    let mut square_periodic_lattice = Lattice::<Ising>::new(row_size*row_size,
+        |site: usize| -> SmallVec<usize> {
+            let mut neighbors = SmallVec::new();
+            if site % row_size != 0 {
+                neighbors.push(site - 1)
+            } else {
+                neighbors.push(site + row_size - 1)
+            }
+            if site % row_size != row_size - 1 {
+                neighbors.push(site + 1)
+            } else {
+                neighbors.push(site - row_size + 1)
+            }
+            if site / row_size != 0 {
+                neighbors.push(site - row_size)
+            } else {
+                neighbors.push(row_size * (row_size-1) + site)
+            }
+            if site / row_size != row_size - 1 {
+                neighbors.push(site + row_size)
+            } else {
+                neighbors.push(site - row_size * (site / row_size))
+            }
+            neighbors
+        }
+    );
+
+    let n_trials = 1000;
+    for beta in [0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2.0] {
+        println!("Beta {}, param {}", beta, square_periodic_lattice.run(beta, n_trials));
+        square_periodic_lattice.zero();
+    }
 }

### 2023-05-26 src/spin.rs

@@ -1,0 +1,100 @@
+use std::ops::AddAssign;
+use std::fmt::Debug;
+
+use rand::random;
+
+const MAX_NEIGHBORS: usize = 8;
+
+#[derive(Debug)]
+pub struct Ising {
+    data: i32,
+}
+
+impl<'a> AddAssign<&'a Self> for Ising {
+    fn add_assign(&mut self, rhs: &'a Self) {
+        self.data += rhs.data;
+    }
+}
+
+impl Spin for Ising {
+    type V = u8;
+    fn dot(&mut self, _vec: &Self::V) -> f64 {
+        self.data as f64
+    }
+    fn flip(&mut self, _vec: &Self::V) {
+        self.data *= -1
+    }
+    fn norm(&self) -> f64 {
+        self.data as f64
+    }
+    fn random_vec() -> Self::V {
+        0
+    }
+    fn start() -> Self {
+        Self { data: if random::<bool>() {1} else {-1} }
+    }
+    fn zero() -> Self {
+        Self { data: 0 }
+    }
+}
+
+
+pub struct SmallVec<T> {
+    data: [Option<T>; MAX_NEIGHBORS]
+}
+
+pub struct SmallVecIter<'a, T> {
+    origin: &'a SmallVec<T>,
+    index: usize,
+}
+
+impl<'a, T> Iterator for SmallVecIter<'a, T> {
+    type Item = &'a T;
+    fn next(&mut self) -> Option<Self::Item> {
+        let val = self.origin.data[self.index].as_ref();
+        self.index += 1;
+        val
+    }
+}
+
+impl<T: Copy> SmallVec<T> {
+    pub fn new() -> Self {
+        Self {
+            data: [None; MAX_NEIGHBORS],
+        }
+    }
+
+    pub fn push(&mut self, val: T) {
+        for i in 0..MAX_NEIGHBORS {
+            if let None = self.data[i] {
+                self.data[i] = Some(val);
+                return;
+            }
+        }
+        panic!("Could not push because the list was full")
+    }
+
+    pub fn iter<'a>(&'a self) -> SmallVecIter<'a, T> {
+        SmallVecIter {
+            origin: &self,
+            index: 0,
+        }
+    }
+
+    pub fn iter_mut<'a>(&'a mut self) -> SmallVecIter<'a, T> {
+        SmallVecIter {
+            origin: self,
+            index: 0,
+        }
+    }
+}
+
[…]

### 2023-05-26 python/plot.py
Confirmed square lattice transition temperature

@@ -1,0 +1,31 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+betas = [0.4, 0.41000000000000003, 0.42000000000000004, 0.43, 0.44, 0.45, 0.46, 0.47000000000000003, 0.48, 0.49, 0.5]
+order_params = [0.051732381184895836, 0.06700331164944556, 0.08913872831611223, 0.15970816663516466, 0.5101541498655914, 0.750468264343918, 0.814727783203125, 0.851683093655494, 0.8773756129767305, 0.8964162847047211, 0.911744640719506]
+susceptibilities = [0.003807657508439915, 0.00633659670429845, 0.013305633298812373, 0.038892643426054266, 0.2958016754478537, 0.5644888724050214, 0.6589932287892988, 0.7255762007928663, 0.7734500310754263, 0.802207290485341, 0.8308364550272623]
+
+betas = np.array(betas)
+mid_betas = (betas[1:] + betas[:-1]) / 2
+order_params = np.array(order_params)
+susceptibilities = np.array(susceptibilities)
+order_param_slope = (order_params[1:] - order_params[:-1])
+susceptibility_slope = (susceptibilities[1:] - susceptibilities[:-1])
+
+fig, (ax_param, ax_sus) = plt.subplots(figsize=(6, 8), nrows=2, sharex=True)
+
+ax_param.scatter(betas, order_params, color="C0")
+ax_param.plot(betas, order_params, color="C0")
+ax_param.scatter(mid_betas, order_param_slope, edgecolor="C0", facecolor='none')
+ax_param.plot(mid_betas, order_param_slope, color="C0", linestyle='dashed')
+ax_param.set_ylabel("Magnetization")
+
+ax_sus.scatter(betas, susceptibilities, color="C1")
+ax_sus.plot(betas, susceptibilities, color="C1")
+ax_sus.scatter(mid_betas, susceptibility_slope, edgecolor="C1", facecolor='none')
+ax_sus.plot(mid_betas, susceptibility_slope, color="C1", linestyle='dashed')
+ax_sus.set_xlabel("beta")
+ax_sus.set_ylabel("Susceptibility")
+
+fig.tight_layout()
+fig.savefig("../figs/params.png")

### 2023-05-26 src/funcs.rs

@@ -1,0 +1,28 @@
+use crate::spin::SmallVec;
+
+pub const SQUARE_ROW_SIZE: usize = 128;
+
+pub fn square(site: usize) -> SmallVec<usize> {
+    let mut neighbors = SmallVec::new();
+    if site % SQUARE_ROW_SIZE != 0 {
+        neighbors.push(site - 1)
+    } else {
+        neighbors.push(site + SQUARE_ROW_SIZE - 1)
+    }
+    if site % SQUARE_ROW_SIZE != SQUARE_ROW_SIZE - 1 {
+        neighbors.push(site + 1)
+    } else {
+        neighbors.push(site - SQUARE_ROW_SIZE + 1)
+    }
+    if site / SQUARE_ROW_SIZE != 0 {
+        neighbors.push(site - SQUARE_ROW_SIZE)
+    } else {
+        neighbors.push(SQUARE_ROW_SIZE * (SQUARE_ROW_SIZE-1) + site)
+    }
+    if site / SQUARE_ROW_SIZE != SQUARE_ROW_SIZE - 1 {
+        neighbors.push(site + SQUARE_ROW_SIZE)
+    } else {
+        neighbors.push(site - SQUARE_ROW_SIZE * (site / SQUARE_ROW_SIZE))
+    }
+    neighbors
+}

### 2023-05-26 src/lattice.rs

@@ -1,11 +1,30 @@
-use std::collections::VecDeque;
+use std::fmt::Display;
 
 use crate::spin::{Spin, SmallVec};
 use rand::random;
 
+const CLUSTER_QUEUE_MAX_SIZE: usize = 10000;
+const BURN_IN: usize = 256;
+const RUNS_PER_TRIAL: usize = 16;
+
+pub struct Report {
+    pub order_parameter: f64,
+    pub order_parameter_err: f64,
+    pub susceptibility: f64,
+    pub susceptibility_err: f64,
+}
+
+impl Display for Report {
+    fn fmt(&self, f: &mut std::fmt::Formatter<'_>) -> std::fmt::Result {
+        writeln!(f, "Order parameter {} +/- {}", self.order_parameter, self.order_parameter_err)?;
+        writeln!(f, "Susceptibility {} +/- {}", self.susceptibility, self.susceptibility_err)
+    }
+}
+
 pub struct Lattice<S: Spin> {
     data: Vec<S>,
     neighbors: Vec<SmallVec<usize>>,
+    cluster_queue: [usize; CLUSTER_QUEUE_MAX_SIZE],
 }
 
 impl<S: Spin> Lattice<S> {
@@ -20,6 +39,7 @@ impl<S: Spin> Lattice<S> {
         Self {
             data,
             neighbors,
+            cluster_queue: [0; CLUSTER_QUEUE_MAX_SIZE],
         }
     }
 
@@ -29,44 +49,69 @@ impl<S: Spin> Lattice<S> {
         }
     }
 
-    pub fn run(&mut self, beta: f64, n_trials: usize) -> f64 {
-        let rolling_avg_count = 500;
-        let runs_per_trial = 16;
-        
-        self.evolve(beta, runs_per_trial * (n_trials - rolling_avg_count));
-        let mut rolling_avg = 0.0;
-        for _ in 0..rolling_avg_count {
-            self.evolve(beta, runs_per_trial);
-            rolling_avg += self.order_param();
+    pub fn run(&mut self, beta: f64, n_trials: usize) -> Report {
+        let num_avgs = n_trials - BURN_IN;
+
+        self.evolve(beta, RUNS_PER_TRIAL * BURN_IN);
+        let mut avg_param = 0.0;
+        let mut avg_sus = 0.0;
+        let mut avg_param2 = 0.0;
+        let mut avg_sus2 = 0.0;
+        for _ in BURN_IN..n_trials {
+            let sus = self.evolve(beta, RUNS_PER_TRIAL);
[…]

### 2023-05-26 src/main.rs

@@ -1,41 +1,44 @@
 #![allow(dead_code)]
 mod lattice;
 mod spin;
+mod funcs;
 use lattice::Lattice;
-use spin::{Ising, SmallVec};
+use spin::Ising;
+use funcs::{square, SQUARE_ROW_SIZE};
+
+fn linspace(start: f64, end: f64, count: usize) -> Vec<f64> {
+    if count == 1 {
+        return vec![start];
+    }
+    let delta = (end - start) / (count - 1) as f64;
+    let mut out = Vec::with_capacity(count);
+    for i in 0..count {
+        out.push(start + delta * i as f64);
+    }
+    out
+}
 
 fn main() {
-    let row_size = 100;
-    let mut square_periodic_lattice = Lattice::<Ising>::new(row_size*row_size,
-        |site: usize| -> SmallVec<usize> {
-            let mut neighbors = SmallVec::new();
-            if site % row_size != 0 {
-                neighbors.push(site - 1)
-            } else {
-                neighbors.push(site + row_size - 1)
-            }
-            if site % row_size != row_size - 1 {
-                neighbors.push(site + 1)
-            } else {
-                neighbors.push(site - row_size + 1)
-            }
-            if site / row_size != 0 {
-                neighbors.push(site - row_size)
-            } else {
-                neighbors.push(row_size * (row_size-1) + site)
-            }
-            if site / row_size != row_size - 1 {
-                neighbors.push(site + row_size)
-            } else {
-                neighbors.push(site - row_size * (site / row_size))
-            }
-            neighbors
-        }
+    // let mut lattice = Lattice::<Ising>::new(
+    //     SQUARE_ROW_SIZE*SQUARE_ROW_SIZE, square
+    // );
+    let mut lattice = Lattice::<Ising>::new(
+        SQUARE_ROW_SIZE*SQUARE_ROW_SIZE, square
     );
 
     let n_trials = 1000;
-    for beta in [0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2.0] {
-        println!("Beta {}, param {}", beta, square_periodic_lattice.run(beta, n_trials));
-        square_periodic_lattice.zero();
+    let mut order_params = Vec::new();
+    let mut susceptibilities = Vec::new();
+    let betas = linspace(0.4, 0.5, 11);
[…]

### 2023-05-26 src/spin.rs

@@ -21,11 +21,11 @@ impl Spin for Ising {
     fn dot(&mut self, _vec: &Self::V) -> f64 {
         self.data as f64
     }
-    fn flip(&mut self, _vec: &Self::V) {
-        self.data *= -1
+    fn flip(&mut self, _vec: &Self::V, success: bool) {
+        self.data *= 1 - 2 * (success as i32)
     }
     fn norm(&self) -> f64 {
-        self.data as f64
+        (self.data as f64).abs()
     }
     fn random_vec() -> Self::V {
         0
@@ -91,7 +91,7 @@ impl<T: Copy> SmallVec<T> {
 
 pub trait Spin: for <'a> AddAssign<&'a Self> + Sized + Debug {
     type V;
-    fn flip(&mut self, vec: &Self::V);
+    fn flip(&mut self, vec: &Self::V, success: bool);
     fn dot(&mut self, vec: &Self::V) -> f64;
     fn random_vec() -> Self::V;
     fn zero() -> Self;

### 2023-06-07 python/phase.py
Correct susceptibility

@@ -1,0 +1,116 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+from scipy.optimize import minimize
+
+WINDOW = (0.25, 0.55)
+
+def load(filename):
+    with open(f"../data/output/{filename}", 'r') as f:
+        betas = f.readline().split(',')[:-1]
+        betas = np.array([float(b) for b in betas])
+        sus = f.readline().split(',')[:-1]
+        sus = np.array([float(s) for s in sus])
+        argsort = np.argsort(betas)
+        betas = betas[argsort]
+        sus = sus[argsort]
+        mid_betas = []
+        sus_deriv = []
+        for i in range(len(sus) - 1):
+            dbeta = betas[i+1] - betas[i]
+            if dbeta == 0: continue
+            dsus = sus[i+1] - sus[i]
+            mid_betas.append((betas[i] + betas[i+1]) / 2)
+            sus_deriv.append(dsus/dbeta)
+
+        mask = (WINDOW[0] < betas) & (betas < WINDOW[1])
+        mask = sus < 0.25
+        sus = sus[mask]
+        betas = betas[mask]
+        return betas, sus, mid_betas, sus_deriv
+    
+def model_fn(param, exponent, scale, xs):
+    tau = param / xs - 1
+    output = (np.abs(tau)**exponent) * scale
+    output[tau > 0] = 0
+    return output
+
+def chisq(params, xs, ys):
+    param, exponent, scale = params
+    return np.sum((ys - model_fn(param, exponent, scale, xs))**2 / 2)
+
+def get_stats_fit(betas, sus):
+    x0 = 0.5, 0.5, 0.5
+    result = minimize(chisq, x0, args=(betas, sus), bounds=(
+        (0.25,0.75),
+        (0,3),
+        (0,1)
+    ))
+
+    fig, ax = plt.subplots()
+    ax.plot(betas, sus)
+    ax.plot(betas, model_fn(result.x[0], result.x[1], result.x[2],betas))
+    ax.plot(betas, model_fn(0.28, 0.5, 1,betas))
+    fig.savefig("test.png")
+    return result.x[0], result.x[1]
+
+def get_stats_steep(betas, sus):
+    max_slope = None
+    max_slope_beta = None
+    max_slope_sus = None
+    if len(sus) <= 1: return np.nan, 1
+    for i in range(len(sus) - 1):
+        dbeta = betas[i+1] - betas[i]
+        if dbeta == 0: continue
[…]

### 2023-06-07 python/plot.py

@@ -1,31 +1,33 @@
 import matplotlib.pyplot as plt
 import numpy as np
 
-betas = [0.4, 0.41000000000000003, 0.42000000000000004, 0.43, 0.44, 0.45, 0.46, 0.47000000000000003, 0.48, 0.49, 0.5]
-order_params = [0.051732381184895836, 0.06700331164944556, 0.08913872831611223, 0.15970816663516466, 0.5101541498655914, 0.750468264343918, 0.814727783203125, 0.851683093655494, 0.8773756129767305, 0.8964162847047211, 0.911744640719506]
-susceptibilities = [0.003807657508439915, 0.00633659670429845, 0.013305633298812373, 0.038892643426054266, 0.2958016754478537, 0.5644888724050214, 0.6589932287892988, 0.7255762007928663, 0.7734500310754263, 0.802207290485341, 0.8308364550272623]
+def load(filename):
+    with open(f"../data/output/{filename}", 'r') as f:
+        betas = f.readline().split(',')[:-1]
+        betas = np.array([float(b) for b in betas])
+        mag = f.readline().split(',')[:-1]
+        mag = np.array([float(s) for s in mag])
+        sus = f.readline().split(',')[:-1]
+        sus = np.array([float(s) for s in sus])
+        argsort = np.argsort(betas)
+        betas = betas[argsort]
+        mag = mag[argsort]
+        sus = sus[argsort]
+        return betas, mag, sus
+    
+betas, mag, sus = load("ising-square.dat")
 
-betas = np.array(betas)
-mid_betas = (betas[1:] + betas[:-1]) / 2
-order_params = np.array(order_params)
-susceptibilities = np.array(susceptibilities)
-order_param_slope = (order_params[1:] - order_params[:-1])
-susceptibility_slope = (susceptibilities[1:] - susceptibilities[:-1])
+fig, (ax_mag, ax_sus) = plt.subplots(figsize=(6, 10), nrows=2, sharex=True)
 
-fig, (ax_param, ax_sus) = plt.subplots(figsize=(6, 8), nrows=2, sharex=True)
-
-ax_param.scatter(betas, order_params, color="C0")
-ax_param.plot(betas, order_params, color="C0")
-ax_param.scatter(mid_betas, order_param_slope, edgecolor="C0", facecolor='none')
[…]

### 2023-06-07 src/funcs.rs

@@ -1,28 +1,113 @@
+use std::io::Read;
+
+use npy::NpyData;
+
 use crate::spin::SmallVec;
 
-pub const SQUARE_ROW_SIZE: usize = 128;
+pub fn square_fn(row_size: usize) -> impl Fn(usize) -> SmallVec<(usize, f32)> {
+    rect_fn(row_size, 1.)
+}
 
-pub fn square(site: usize) -> SmallVec<usize> {
-    let mut neighbors = SmallVec::new();
-    if site % SQUARE_ROW_SIZE != 0 {
-        neighbors.push(site - 1)
-    } else {
-        neighbors.push(site + SQUARE_ROW_SIZE - 1)
+pub fn rect_fn(row_size: usize, t2: f32) -> impl Fn(usize) -> SmallVec<(usize, f32)> {
+    move |site: usize| {
+        let mut neighbors = SmallVec::new();
+        if site % row_size != 0 {
+            neighbors.push((site - 1, 1.))
+        } else {
+            neighbors.push((site + row_size - 1, 1.))
+        }
+        if site % row_size != row_size - 1 {
+            neighbors.push((site + 1, 1.))
+        } else {
+            neighbors.push((site - row_size + 1, 1.))
+        }
+        if site / row_size != 0 {
+            neighbors.push((site - row_size, t2))
+        } else {
+            neighbors.push((row_size * (row_size-1) + site, t2))
+        }
+        if site / row_size != row_size - 1 {
+            neighbors.push((site + row_size, t2))
+        } else {
+            neighbors.push((site - row_size * (site / row_size), t2))
+        }
+        neighbors
     }
-    if site % SQUARE_ROW_SIZE != SQUARE_ROW_SIZE - 1 {
-        neighbors.push(site + 1)
-    } else {
-        neighbors.push(site - SQUARE_ROW_SIZE + 1)
-    }
-    if site / SQUARE_ROW_SIZE != 0 {
-        neighbors.push(site - SQUARE_ROW_SIZE)
-    } else {
-        neighbors.push(SQUARE_ROW_SIZE * (SQUARE_ROW_SIZE-1) + site)
-    }
-    if site / SQUARE_ROW_SIZE != SQUARE_ROW_SIZE - 1 {
-        neighbors.push(site + SQUARE_ROW_SIZE)
-    } else {
-        neighbors.push(site - SQUARE_ROW_SIZE * (site / SQUARE_ROW_SIZE))
[…]

### 2023-06-07 src/lattice.rs

@@ -3,32 +3,31 @@ use std::fmt::Display;
 use crate::spin::{Spin, SmallVec};
 use rand::random;
 
-const CLUSTER_QUEUE_MAX_SIZE: usize = 10000;
+const CLUSTER_QUEUE_MAX_SIZE: usize = 256_000;
 const BURN_IN: usize = 256;
 const RUNS_PER_TRIAL: usize = 16;
 
+/// Result from a complete run of a lattice
 pub struct Report {
-    pub order_parameter: f64,
-    pub order_parameter_err: f64,
-    pub susceptibility: f64,
-    pub susceptibility_err: f64,
+    pub magnetization: f32,
+    pub susceptibility: f32,
 }
 
 impl Display for Report {
     fn fmt(&self, f: &mut std::fmt::Formatter<'_>) -> std::fmt::Result {
-        writeln!(f, "Order parameter {} +/- {}", self.order_parameter, self.order_parameter_err)?;
-        writeln!(f, "Susceptibility {} +/- {}", self.susceptibility, self.susceptibility_err)
+        writeln!(f, "Magnetization {}", self.magnetization)?;
+        writeln!(f, "Susceptibility {}", self.susceptibility)
     }
 }
 
 pub struct Lattice<S: Spin> {
     data: Vec<S>,
-    neighbors: Vec<SmallVec<usize>>,
+    neighbors: Vec<SmallVec<(usize, f32)>>,
     cluster_queue: [usize; CLUSTER_QUEUE_MAX_SIZE],
 }
 
 impl<S: Spin> Lattice<S> {
-    pub fn new(num_sites: usize, neighbor_func: impl Fn(usize) -> SmallVec<usize>) -> Self {
+    pub fn new(num_sites: usize, neighbor_func: impl Fn(usize) -> SmallVec<(usize, f32)>) -> Self {
         let mut data = Vec::with_capacity(num_sites);
         let mut neighbors = Vec::with_capacity(num_sites);
         for site_index in 0..num_sites {
@@ -49,35 +48,35 @@ impl<S: Spin> Lattice<S> {
         }
     }
 
-    pub fn run(&mut self, beta: f64, n_trials: usize) -> Report {
-        let num_avgs = n_trials - BURN_IN;
+    pub fn run(&mut self, beta: f32, n_trials: usize, checkpoint: (usize, f32)) -> Report {
+        let mut num_avgs = n_trials - BURN_IN;
 
[…]

### 2023-06-07 src/main.rs

@@ -1,44 +1,236 @@
 #![allow(dead_code)]
+#![allow(unused_imports)]
 mod lattice;
 mod spin;
 mod funcs;
-use lattice::Lattice;
-use spin::Ising;
-use funcs::{square, SQUARE_ROW_SIZE};
+use core::num;
+use std::{fs::File, io::Write, thread, sync::Arc};
 
-fn linspace(start: f64, end: f64, count: usize) -> Vec<f64> {
+use lattice::Lattice;
+use spin::{Ising, Spin};
+use funcs::{square_fn, load_penrose, load_einstein, rect_fn};
+
+use crate::spin::XY;
+
+const NUM_THREADS: usize = 8;
+
+struct Data {
+    betas: Vec<f32>,
+    magnetizations: Vec<f32>,
+    susceptibilities: Vec<f32>,
+}
+
+impl Data {
+    fn save(&self, filename: &str) {
+        let mut f = File::create(format!("data/output/{}.dat", filename)).unwrap();
+        for entry in &self.betas {
+            write!(&mut f, "{},", entry).unwrap();
+        }
+        writeln!(f,"").unwrap();
+        for entry in &self.magnetizations {
+            write!(&mut f, "{},", entry).unwrap();
+        }
+        writeln!(f,"").unwrap();
+        for entry in &self.susceptibilities {
+            write!(&mut f, "{},", entry).unwrap();
+        }
+        writeln!(f,"").unwrap();
+    }
+}
+
+fn linspace(start: f32, end: f32, count: usize) -> Vec<f32> {
     if count == 1 {
         return vec![start];
     }
-    let delta = (end - start) / (count - 1) as f64;
+    let delta = (end - start) / (count - 1) as f32;
     let mut out = Vec::with_capacity(count);
     for i in 0..count {
-        out.push(start + delta * i as f64);
+        out.push(start + delta * i as f32);
     }
     out
 }
 
-fn main() {
-    // let mut lattice = Lattice::<Ising>::new(
-    //     SQUARE_ROW_SIZE*SQUARE_ROW_SIZE, square
-    // );
-    let mut lattice = Lattice::<Ising>::new(
-        SQUARE_ROW_SIZE*SQUARE_ROW_SIZE, square
-    );
+fn linspace_ex(start: f32, end: f32, count: usize) -> Vec<f32> {
+    if count == 1 {
+        return vec![start];
+    }
+    let delta = (end - start) / (count + 1) as f32;
[…]

### 2023-06-07 src/spin.rs

@@ -1,13 +1,20 @@
 use std::ops::AddAssign;
 use std::fmt::Debug;
 
-use rand::random;
+use rand::prelude::*;
+use rand_distr::StandardNormal;
 
-const MAX_NEIGHBORS: usize = 8;
+const MAX_NEIGHBORS: usize = 12;
 
 #[derive(Debug)]
 pub struct Ising {
     data: i32,
+}
+
+#[derive(Debug)]
+pub struct XY {
+    x: f32,
+    y: f32,
 }
 
 impl<'a> AddAssign<&'a Self> for Ising {
@@ -16,16 +23,23 @@ impl<'a> AddAssign<&'a Self> for Ising {
     }
 }
 
+impl<'a> AddAssign<&'a Self> for XY {
+    fn add_assign(&mut self, rhs: &'a Self) {
+        self.x += rhs.x;
+        self.y += rhs.y;
+    }
+}
+
 impl Spin for Ising {
     type V = u8;
-    fn dot(&mut self, _vec: &Self::V) -> f64 {
-        self.data as f64
+    fn dot(&mut self, _vec: &Self::V) -> f32 {
+        self.data as f32
     }
     fn flip(&mut self, _vec: &Self::V, success: bool) {
         self.data *= 1 - 2 * (success as i32)
     }
-    fn norm(&self) -> f64 {
-        (self.data as f64).abs()
+    fn norm(&self) -> f32 {
+        (self.data as f32).abs()
     }
     fn random_vec() -> Self::V {
         0
@@ -38,7 +52,38 @@ impl Spin for Ising {
     }
 }
 
+impl Spin for XY {
+    type V = (f32, f32);
+    fn dot(&mut self, vec: &Self::V) -> f32 {
+        self.x * vec.0 + self.y * vec.1
+    }
+    fn flip(&mut self, vec: &Self::V, success: bool) {
+        let dot = (vec.0 * self.x + vec.1 * self.y) * (success as i32 as f32);
+        self.x -= 2. * dot * vec.0;
+        self.y -= 2. * dot * vec.1;
+        let norm = 1./(self.x*self.x + self.y*self.y).sqrt();
+        self.x *= norm;
+        self.y *= norm;
+    }
+    fn norm(&self) -> f32 {
+        (self.x * self.x + self.y * self.y).sqrt()
+    }
+    fn random_vec() -> Self::V {
+        let x = 0.1 * thread_rng().sample::<f32,_>(StandardNormal);
[…]

### 2023-06-08 paper/main.tex
Better plots and fixed up the data

@@
-\documentclass[11pt]{revtex4}
+\documentclass[amsmath,amssymb,aps,twocolumn]{revtex4-2}
 
+\usepackage{bm}
+\usepackage{graphicx}
 \usepackage[colorlinks=true]{hyperref}
+
+\newcommand{\braket}[1]{\left \langle #1 \right \rangle}
+\newcommand{\parens}[1]{\left ( #1 \right )}
+\newcommand{\jtd}[1]{{\color{red}\textbf{JTD: #1}}}
 
@@
 \begin{abstract}
-  Abstract
+  A shape known as the ``einstein tile'' has been discovered which tiles the plane aperiodically.
+We define the Ising, XY, and transverse Ising models on the einstein tiling and compute the magnetization and susceptibility of the tiling as a function of temperature using Monte Carlo methods.
+This allows us to produce phase diagrams of the relationship between the critical temperature of the tiling and the dimensions of the einstein tile.
 \end{abstract}
@@
 
-Hello there.
+\section{Phase Transitions and Spin Models}
+The Ising model was one of the first statistical mechanics models to contain a phase transition, and remains one of the only exactly solvable examples.
+It consists of defining a variable $s_i = \pm 1$ at every site of a lattice with $N$ sites as a classical approximation of the up-or-down spin of a fermion.
+On further analogy with fermionic spins, it asserts that the spins interact by defining the classical Hamiltonian
+\begin{equation}
+  H = -J \sum_{\braket{ij}} s_i s_j + h \sum_i s_i
+  \label{eqn:ising}
+\end{equation}
+where $h$ is an applied magnetic field. This model's phase transition is most visible in the magnetization, defined as
+\begin{equation}
+  M = \frac{1}{N}\braket{\sum_{i} s_i}
+  \label{eqn:magnetization}
+\end{equation}
+where the brackets indicate an average over the statistical ensemble of states, weighted by the Boltzmann factor $e^{-\beta H}$. Another interesting parameter is the susceptibility
+\begin{equation}
+  \chi = \frac{\partial M}{\partial h} = \frac{1}{N} \braket{\sum_{i} (s_i - M)^2}.
+\end{equation}
+Now that the susceptibility has been defined, we will set $h$ to zero and study the Ising model in the zero magnetic field limit.
+
+The reason why a phase transition occurs is as follows.
+The Hamiltonian displays $s_i\rightarrow -s_i$ symmetry when when $h=0$.
+If the thermodynamic system explores all of phase space, then the average thermodynamic state should possess the same symmetries of the Hamiltonian so that $M=0$.
+However, for low temperature we also expect the system to attain the lowest possible energy.
+This state consists of all spins up (or all spins down) has $M = 1$ (or $-1$), which violates the expectation of no magnetization.
+The resolution to this apparent contradiction that there exists a critical temperature $T_c$ below which energetic considerations dominate and the crystal is magnetic ($|M|>0$), and above which the symmetry of phase space dominates and the crystal is not magnetic ($M = 0$).
+
+The phenomenon of critical temperature has been generalized to other theories.
+In particular, if the system's variables are arranged as a members of a field $s(\bm x)$, then the free energy automatically comes in a form that allows phase transitions using a method known as Landau field theory.
+Taking the Ising model as an example, we could approximate the $s_i$ spins defined on discrete points $x_i$ as a field $s(\bm x)$ defined for any position $\bm x$.
+The original Hamiltonian contained translational, rotational, and $s\rightarrow-s$ symmetry, and therefore the free energy should as well.
+Expanding it in powers of $s(\bm x)$ guarantees the form
+\begin{equation}
+F \propto -\int d^d x \, \parens{(\nabla s(\bm x))^2 + A s(\bm x)^2 + B s(\bm x)^4 + \dots}
+\label{eqn:landau}
+\end{equation}
+where $d$ is the number of dimensions of the lattice and $A$ and $B$ are arbitrary constants.
+If $A>0$, then free energy is minimized at every point by setting $s(\bm x)$.
+However, if $A<0$, then minimizing the free energy necessitates setting $s(\bm x)$ to some nonzero value, thereby magnetizing the crystal and creating a phase transition.
+In principle, all the parameters of the model including $A$ may depend on temperature.
+All that is necessary to describe a phase transition is therefore to write a form of $A$ which allows it to cross from positive to negative at a critical temperature $T_c$.
+
+The utility of this field theory approximation is in its generality.
+The Ising model and all models like it are Landau field theories and are therefore described by the sameEq.~\ref{eqn:landau}.
+Therefore, all Ising models have the same near-critical-point behavior regardless of the lattice on which they are defined.
+An Ising model's only unique properties are their values of $A$ and $B$; i.e., the value of the critical temperature.
+
+In addition to the Ising model, we study the XY model which consists of promoting $s_i$ to a two-dimensional unit vector defined at every lattice point.
+The interaction term in the Hamiltonian becomes $\bm s_i \cdot \bm s_j$.
+This XY model has different $T_c$ than the Ising model and can even contain vortices, which are topologically protected quasiparticles where the curl of $\bm s_i$ is nonzero.
+A third spin model is the Transverse Ising Model (TIM), which consists of replacing $s_i$ in Eq.~\ref{eqn:ising} with the Pauli matrices $\bm \sigma_i$ and $\bm \sigma_j$ and interpreting the result as a quantum Hamiltonian.
+Past research \jtd{citation needed} has shown that the $d$-dimensional TIM is equivalent to a classical Ising model in $d+1$ dimensions in the limit of a very strong interaction strength in the extra dimension.
+Thus, critical temperatures for all three of these models can be found using the same techniques.
+
+In this final project, we will use Monte Carlo methods discussed in the next section to evaluate the critical temperature of these three models for the recently discovered einstein tiling (figure \ref{fig:einstein}).
+The einstein tile (German for ``one stone'') is the first shape known to tile the plane in a non-repeating pattern \cite{smith2023aperiodic}.
+Such aperiodic tilings are called quasicrystals, and a two-tile version known as the Penrose tiling has gained much traction in the literature \jtd{mention its critical points.
+Also show the einstein tile}.
+An interesting fact about the einstein tile is that its dimensions are adjustable.
+The tile consists of two side lengths $a$ and $b$, and any ratio of $a/b$ leads to an aperiodic tiling except $a/b=0$, 1, or $\infty$.
+If we treat the spin-spin interaction as weaker across the longer bonds, then this $a/b$ ratio represents a free parameter of the crystal, so that the critical temperature is a function of $a/b$.
+Generally, we take the interaction strength $J$ across a bond of length $\ell$ to be $J = J_0 \ell^{-2}$ where $J_0$ is constant.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{../figs/einstein.pdf}
+  \caption{The einstein tiling}
+  \label{fig:einstein}
+\end{figure}
+
+\section{Cluster Monte Carlo Methods: the Wolff Algorithm}
+Solving the Ising model is generally difficult.
+Ising himself solved the Ising model for a one-dimensional regular crystal, and a solution was produced 20 years later for a two dimensional square lattice.
+The critical temperature can sometimes be worked out by a clever trick for the Ising model when the lattice is self-dual, such as the square lattice \jtd{cite}.
+Even when the lattice is not self-dual, if the lattice shares enough properties with its dual the connection can be used to determine the critical temperature \jtd{cite}.
+
+However, the einstein lattice (like many lattices) satisfies none of these properties, so that we are reduced to numerically computing the ensemble average of Eq.~\ref{eqn:magnetization}.
+This average is extremely expensive to compute because the number of states in the ensemble is $2^N$ for the Ising model.
+However, approximate results can be obtained by computing the sum over a much smaller, random selection of states $K$.
+This approximation technique is called a Monte Carlo method.
+
+The method of selecting states for $K$ is critical to the accuracy of a Monte Carlo method.
+It needs to add states $\psi$ to $K$ with some probability $P(\psi)$ which prefers states that are representative of the full ensemble, otherwise the convergence of the ensemble average will be poor.
+Many states are not representative, making it difficult to find a new state $\psi$ which has high enough $P(\psi)$ to be added to $K$.
+One way to resolve this problem is to make $\psi$ very similar to a state $\phi$ already in $K$.
+Assuming $\phi$ is representative of the ensemble, then $\psi$ is also likely to be represented by the ensemble.
+We then add $\psi$ to $K$ with some probability $P(\phi \rightarrow \psi)$, which is a function we discuss in the next paragraph.
+The technique of using previous states $\phi$ to create a new state $\psi$ is called a Markov Chain Monte Carlo method (MCMC).
+
+For an MCMC to be correct, $P(\phi \rightarrow \psi)$ must be reversible, meaning that the probability to go from $\phi$ to $\psi$ must be equal to the probability to go from $\psi$ to $\phi$.
+This condition is necessary to prevent $K$ from getting stuck in an unrepresentative region of phase space.
+The probability to go from $\phi$ to $\psi$ is the probability for the MCMC to be at state $\phi$ in the first place times the probability to switch to $\psi$: $P(\phi \rightarrow \psi)P(\phi)$.
+Thus, detailed balance requires
+\begin{equation}
+  P(\phi \rightarrow \psi)P(\phi) = P(\psi \rightarrow \phi)P(\psi)
+  \label{eqn:detailed-balance}
+\end{equation}
+Any method of generating a new state $\psi$ from the old state $\phi$ that satisfies Eq.~\ref{eqn:detailed-balance} is a valid MCMC.
+
+There are many methods of generating new states, but the one we will use in this paper is the Wolff algorithm, which is known to converge well near the critical temperature \cite{wolff1989collective}.
+The algorithm generates a new state $\psi$ by probabilistically choosing a cluster of spins and then flipping all the spins at once.
+The cluster starts the cluster at a random location $i$.
+It then adds each neighbor $j$ to the cluster with probability
+\begin{equation}
+  P(i,j) = 1 - \exp\parens{\mathrm{min}\left[0, 2\beta s_i s_j \right]}.
+  \label{eqn:wolff-algo}
+\end{equation}
+For every spin $j$ which was added to the cluster, it then uses the same equation to assign $j$'s neighbors to the cluster until every neighbor has been rejected.
+At that point, the cluster's spins are flipped to form the new state $\psi$.
+Ref.~\cite{wolff1989collective} confirms that Eq.~\ref{eqn:wolff-algo} satisfies the detailed balance constraint and therefore is a good MCMC.
+
+For the XY model, Eq.~\ref{eqn:wolff-algo} must be modified to handle vector-valued spins. The approach is to choose a random vector $\bm r$ as a direction against which to compare to spins, so that
+\begin{equation}
+  P(i,j) = 1 - \exp\parens{\mathrm{min}\left[0, 2\beta J (\bm r \cdot \bm s_i)(\bm r \cdot \bm s_j) \right]}.
+\end{equation}
+For consistency, we keep $\bm r$ the same throughout the cluster, but for the next cluster we re-draw a new $\bm r$.
+
+A third and final extension is necessary to the Wolff algorithm in order to handle quantum spin models.
+As mentioned above, the TIM in two dimensions is equivalent to the Ising model in three dimensions where interaction strengths $J$ are very strong in the new direction.
+We could therefore stack many two-dimensional crystals to form a three-dimensional crystal on which we run the Wolff algorithm, but this is inefficient because the large $J$ along the third axis will make the clusters grow large in this direction.
+Ref.~\cite{blote2002cluster} speeds up the process by analytically computing the behavior of the Wolff algorithm along the third axis in the limit of large $J$.
+The classical Wolff algorithm can then be run on the two-dimensional lattice while the spins on the third axis are set according to this analytical result.
+
+\section{Results}
+
+The upper left panel of Figure \ref{fig:square} shows the magnetization $M$ and susceptibility $\chi$ of the Ising, XY, and TI models executed on a square lattice as a function of the lattice width $L = \sqrt{N}$.
+The critical temperature of the Ising model on this lattice is known to be $T_c = \frac{2}{\ln (1 + \sqrt{2})} \approx 2.27$ in units where $k_B = J = 1$ due to the Onsager solution, and indeed the magnetization falls sharpest near this location.
+For the XY model, the critical temperature is approximately $0.8816$, which is likewise close to the drop in $M$.
+The location of these transitions is not perfect, and $M$ it is not strictly zero for $T<T_c$ due to finite-size effects.
+Larger lattice sizes lead to smaller $M$ for $T<T_c$.
+In the XY model the critical temperature is substantially lower due to the increased degrees of freedom of the spin system.
+Furthermore, $M$ approaches its low-temperature limit of 1 linearly for low temperature rather than exponentially, which is a qualitative difference between the Ising and XY models.
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=0.49\linewidth]{../figs/square.png}\hfill
+  \includegraphics[width=0.49\linewidth]{../figs/square.png}
+  \caption{Magnetization and susceptibility as a function of $T_c$ under the XY and Ising models.
+\textit{Left}: A square lattice with $N=128^2=16,384$ sites.
+\textit{Right}: A Penrose lattice with $N=21,316$ sites.}
+  \label{fig:square}
+\end{figure*}
+
+The susceptibility in the bottom left panel produces a maximum at $T_c$, decaying to zero more rapidly on the low-temperature side than high-temperature.
+For an infinite lattice, the peak will be likewise infinite, but it is blunted by finite $N$.
+The peak in susceptibility can be used as a measure of the location of $T_c$, as can the point of maximum slope of $M$.
+
+In both crystals and both models, the point of maximum susceptibility and maximum $dM/dT$ occur at approximately the same location.
+For the sake of approximating the critical temperature, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects the $M=1$ axis.
+This approximation results in $T_c\approx 2.26$ for the Ising model on a square lattice and $T_c\approx 0.922$, within a few percent of the true result.
+
+To demonstrate that the $M$-$T$ relationship is not strongly dependent on the lattice, we also compute the same quantities for the Penrose lattice, shown in the right two panels of figure \ref{fig:square}.
+We see the same behavior in how $M$ and $\chi$ approach $T=0$ and $\infty$.
+The behavior near $T_c$ is slightly different in the steepness of the curves due to the difference in the number of sites.
+The major difference between the two is the actual value of $T_c$, which is slightly higher for Penrose.
+\jtd{Why?
+State TC and figure out a way to put errors on it?
+Also draw TC on the plot.}
+
+The square lattice can be stretched into a rectangular lattice by allowing the horizontal and vertical bonds to differ in length.
+Assigning different interaction strengths $J_1$ and $J_2$ to the two bond lengths gives a free parameter $J_1/J_2$.
+Figure \ref{fig:ising-rect} shows the critical temperature as a function of $J_1/J_2$ where we have fixed the tile area to be constant: $J_1J_2 = 1$.
+
+A similar procedure shows the phase diagram for the einstein tiling, with the ratio of side lengths $a/b$ as the free parameter
+
+\section{Conclusion}
+
+\bibliography{cmt.bib}
 

### 2023-06-08 python/phase.py

@@ -1,116 +1,51 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import os
-from scipy.optimize import minimize
+from util import load, get_stats_steep
 
 WINDOW = (0.25, 0.55)
 
-def load(filename):
-    with open(f"../data/output/{filename}", 'r') as f:
-        betas = f.readline().split(',')[:-1]
-        betas = np.array([float(b) for b in betas])
-        sus = f.readline().split(',')[:-1]
-        sus = np.array([float(s) for s in sus])
-        argsort = np.argsort(betas)
-        betas = betas[argsort]
-        sus = sus[argsort]
-        mid_betas = []
-        sus_deriv = []
-        for i in range(len(sus) - 1):
-            dbeta = betas[i+1] - betas[i]
-            if dbeta == 0: continue
-            dsus = sus[i+1] - sus[i]
-            mid_betas.append((betas[i] + betas[i+1]) / 2)
-            sus_deriv.append(dsus/dbeta)
 
-        mask = (WINDOW[0] < betas) & (betas < WINDOW[1])
-        mask = sus < 0.25
-        sus = sus[mask]
-        betas = betas[mask]
-        return betas, sus, mid_betas, sus_deriv
-    
-def model_fn(param, exponent, scale, xs):
-    tau = param / xs - 1
-    output = (np.abs(tau)**exponent) * scale
-    output[tau > 0] = 0
-    return output
-
-def chisq(params, xs, ys):
-    param, exponent, scale = params
-    return np.sum((ys - model_fn(param, exponent, scale, xs))**2 / 2)
-
-def get_stats_fit(betas, sus):
-    x0 = 0.5, 0.5, 0.5
-    result = minimize(chisq, x0, args=(betas, sus), bounds=(
-        (0.25,0.75),
-        (0,3),
-        (0,1)
-    ))
-
-    fig, ax = plt.subplots()
-    ax.plot(betas, sus)
-    ax.plot(betas, model_fn(result.x[0], result.x[1], result.x[2],betas))
-    ax.plot(betas, model_fn(0.28, 0.5, 1,betas))
-    fig.savefig("test.png")
-    return result.x[0], result.x[1]
-
-def get_stats_steep(betas, sus):
-    max_slope = None
-    max_slope_beta = None
-    max_slope_sus = None
-    if len(sus) <= 1: return np.nan, 1
-    for i in range(len(sus) - 1):
-        dbeta = betas[i+1] - betas[i]
[…]

### 2023-06-08 python/plot.py

@@ -1,33 +1,57 @@
 import matplotlib.pyplot as plt
 import numpy as np
+from util import load
 
-def load(filename):
-    with open(f"../data/output/{filename}", 'r') as f:
-        betas = f.readline().split(',')[:-1]
-        betas = np.array([float(b) for b in betas])
-        mag = f.readline().split(',')[:-1]
-        mag = np.array([float(s) for s in mag])
-        sus = f.readline().split(',')[:-1]
-        sus = np.array([float(s) for s in sus])
-        argsort = np.argsort(betas)
-        betas = betas[argsort]
-        mag = mag[argsort]
-        sus = sus[argsort]
-        return betas, mag, sus
+# plt.rcParams["text.usetex"] = True
+plt.rcParams["font.size"] = 12
+plt.rcParams["font.family"] = "Courier"
     
-betas, mag, sus = load("ising-square.dat")
+fig, (ax_mag, ax_sus) = plt.subplots(figsize=(6, 7), nrows=2, sharex=True)
+lines = [
+    ("ising-square.dat", "crimson", "dodgerblue", "o"),
+    ("xy-square.dat", "lightsalmon", "cyan", "^"),
+]
+for filename, red, blue, marker in lines:
+    betas, mag, sus = load(filename)
+    ax_mag.plot(1/betas, mag, color=blue, zorder=-10)
+    ax_mag.scatter(1/betas, mag, facecolor=blue, edgecolor='k', marker=marker)
+    ax_sus.plot(1/betas, sus, color=red, zorder=-10)
+    ax_sus.scatter(1/betas, sus, facecolor=red, edgecolor='k', marker=marker)
+ax_sus.set_xlabel("Critical temperature $T_c$")
+ax_mag.set_ylabel("Magnetization $M$")
+ax_sus.set_ylabel("Susceptibility $\chi$")
+ax_sus.set_ylim(1e-1,None)
+ax_mag.set_ylim(0,1)
+ax_sus.set_yscale('log')
+ax_sus.set_xlim(1/betas[-1], 1/betas[0])
+ax_sus.plot([],[],color='k', marker='o', markerfacecolor='lightgray', markeredgecolor='k',label="Ising model")
+ax_sus.plot([],[],color='k', marker='^', markerfacecolor='lightgray', markeredgecolor='k',label="XY model")
[…]

### 2023-06-08 python/util.py

@@ -1,0 +1,32 @@
+import numpy as np
+
+def load(filename):
+    with open(f"../data/output/{filename}", 'r') as f:
+        betas = f.readline().split(',')[:-1]
+        betas = np.array([float(b) for b in betas])
+        mag = f.readline().split(',')[:-1]
+        mag = np.array([float(s) for s in mag])
+        sus = f.readline().split(',')[:-1]
+        sus = np.array([float(s) for s in sus])
+        argsort = np.argsort(betas)
+        betas = betas[argsort]
+        mag = mag[argsort]
+        sus = sus[argsort]
+        return betas, mag, sus
+    
+def get_stats_steep(betas, mag):
+    max_slope = None
+    max_slope_beta = None
+    max_slope_mag= None
+    temps = 1/betas
+    if len(mag) <= 1: return np.nan, 1
+    for i in range(len(mag) - 1):
+        dtemp = temps[i+1] - temps[i]
+        if dtemp == 0: continue
+        dmag = mag[i+1] - mag[i]
+        if max_slope is None or dmag/dtemp < max_slope:
+            max_slope = dmag/dtemp
+            max_slope_mag = (mag[i] + mag[i+1]) / 2
+            max_slope_temp = (temps[i] + temps[i+1]) / 2
+    intercept = max_slope_temp + (0.5 - max_slope_mag) / max_slope
+    return intercept

### 2023-06-08 src/funcs.rs

@@ -40,9 +40,12 @@ pub fn load_penrose(level: usize) -> (usize, impl Fn(usize) ->SmallVec<(usize, f
     std::fs::File::open(format!("data/penrose-{}.npy", level)).unwrap()
         .read_to_end(&mut buf).unwrap();
     let bond_indices = NpyData::<i64>::from_bytes(&buf).unwrap().to_vec();
-
+    let mut max_bond_index = 0;
+    for i in &bond_indices {
+        max_bond_index = max_bond_index.max(*i);
+    }
     (
-        bond_indices.len() as usize / 2,
+        max_bond_index as usize + 1,
         move |site: usize| {
             let mut neighbors = SmallVec::new();
             for pair in bond_indices.chunks(2) {

### 2023-06-08 src/lattice.rs

@@ -4,7 +4,7 @@ use crate::spin::{Spin, SmallVec};
 use rand::random;
 
 const CLUSTER_QUEUE_MAX_SIZE: usize = 256_000;
-const BURN_IN: usize = 256;
+const BURN_IN: usize = 128;
 const RUNS_PER_TRIAL: usize = 16;
 
 /// Result from a complete run of a lattice
@@ -27,7 +27,7 @@ pub struct Lattice<S: Spin> {
 }
 
 impl<S: Spin> Lattice<S> {
-    pub fn new(num_sites: usize, neighbor_func: impl Fn(usize) -> SmallVec<(usize, f32)>) -> Self {
+    pub fn new(num_sites: usize, neighbor_func: &impl Fn(usize) -> SmallVec<(usize, f32)>) -> Self {
         let mut data = Vec::with_capacity(num_sites);
         let mut neighbors = Vec::with_capacity(num_sites);
         for site_index in 0..num_sites {

### 2023-06-08 src/main.rs

@@ -50,6 +50,10 @@ fn linspace(start: f32, end: f32, count: usize) -> Vec<f32> {
     out
 }
 
+fn reciprocal_linspace(start: f32, end: f32, count: usize) -> Vec<f32> {
+    linspace(end, start, count).into_iter().map(|f|{1./f}).collect::<Vec<_>>()
+}
+
 fn linspace_ex(start: f32, end: f32, count: usize) -> Vec<f32> {
     if count == 1 {
         return vec![start];
@@ -60,6 +64,23 @@ fn linspace_ex(start: f32, end: f32, count: usize) -> Vec<f32> {
         out.push(start + delta * (i + 1) as f32);
     }
     out
+}
+
+fn one_pass<S: Spin>(lattice: &mut Lattice<S>, start_temp: f32, end_temp: f32,
+    n_trials: usize, num_betas: usize) -> Data {
+
+    let mut susceptibilities = Vec::new();
+    let mut magnetizations = Vec::new();
+    let betas = reciprocal_linspace(start_temp, end_temp, num_betas);
+    for beta in &betas {
+        lattice.zero();
+        let result = lattice.run(*beta, n_trials, (n_trials / 10, 0.5));
+        magnetizations.push(result.magnetization);
+        susceptibilities.push(result.susceptibility);
+    }
+    Data {
+        betas, magnetizations, susceptibilities
+    }
 }
 
 fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
@@ -73,7 +94,6 @@ fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
     for iter_count in 0..layers {
         let mut new_start = start;
         let mut new_end = end;
-        // println!("{}, {}", start, end);
         let beta_line = if iter_count == 0 {
             linspace(start, end, count_per_iteration)
         } else {
@@ -93,7 +113,6 @@ fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
                 new_end = new_end.min(beta);
             }
 
-            // println!("..{}, {}, {}", beta, result.magnetization, result.susceptibility);
         }
         start = new_start;
         end = new_end;
@@ -105,27 +124,44 @@ fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
 }
 
 fn main() {
-    xy_rect();
-    xy_einstein();
-    ising_rect();
[…]

### 2023-06-09 paper/main.tex
Debugging quantum

@@
-\documentclass[amsmath,amssymb,aps,twocolumn]{revtex4-2}
+\documentclass[amsmath,amssymb,aps,twocolumn,nofootinbib]{revtex4-2}
 
@@
 The phenomenon of critical temperature has been generalized to other theories.
-In particular, if the system's variables are arranged as a members of a field $s(\bm x)$, then the free energy automatically comes in a form that allows phase transitions using a method known as Landau field theory.
-Taking the Ising model as an example, we could approximate the $s_i$ spins defined on discrete points $x_i$ as a field $s(\bm x)$ defined for any position $\bm x$.
+In particular, if the system's variables are arranged as a members of a field $s(\bm x)$, then the free energy may be approximated as an integral over functions of $s(\bm x)$.
+This approximation is called Landau Field Theory and it yields a form which is readily able to create phase transitions.
+Taking the Ising model as an example, we could approximate the $s_i$ spins defined on discrete points $x_i$ as the field $s(\bm x)$ defined for any position $\bm x$.
 The original Hamiltonian contained translational, rotational, and $s\rightarrow-s$ symmetry, and therefore the free energy should as well.
@@
 The utility of this field theory approximation is in its generality.
-The Ising model and all models like it are Landau field theories and are therefore described by the sameEq.~\ref{eqn:landau}.
-Therefore, all Ising models have the same near-critical-point behavior regardless of the lattice on which they are defined.
-An Ising model's only unique properties are their values of $A$ and $B$; i.e., the value of the critical temperature.
+All Ising-like models like it are Landau field theories and are all described by Eq~\ref{eqn:landau}, meaning they have the same near-critical-point behavior.
+This statement is independent of the lattice or interaction strengths; it depends only on the field $\bm s$ and the dimensionality of the lattice.
+A particular lattice's only unique properties its its values of $A$ and $B$; i.e., the value of its critical temperature.
 
@@
 
+We implemented the Wolff algorithm with a custom code base \footnote{\jtd{LINK}} and executed for a variety of spin models and crystals.
+Except when stated otherwise, we used a finite-size lattice with initial conditions of random spins.
+Magnetization $M$ and susceptibility $\chi$ measurements were collected every 16 cluster flips to prevent autocorrelation.
+10,000 $M$ and $\chi$ measurements were made when $M$ was small, and when $M$ was large we made 1,000 measurements because the algorithm was found to be more stable for $T< T_c$.
+
 The upper left panel of Figure \ref{fig:square} shows the magnetization $M$ and susceptibility $\chi$ of the Ising, XY, and TI models executed on a square lattice as a function of the lattice width $L = \sqrt{N}$.
@@
   \includegraphics[width=0.49\linewidth]{../figs/square.png}\hfill
-  \includegraphics[width=0.49\linewidth]{../figs/square.png}
+  \includegraphics[width=0.49\linewidth]{../figs/penrose.png}
   \caption{Magnetization and susceptibility as a function of $T_c$ under the XY and Ising models.

### 2023-06-09 python/phase.py

@@ -8,26 +8,26 @@ WINDOW = (0.25, 0.55)
 
 def get_phase(start_name):
     t2s = []
-    beta_criticals = []
+    temp_criticals = []
     for f in os.listdir("../data/output/"):
         if f.startswith(start_name):
             t2 = float(f[len(start_name):-4])
             betas, mag, sus = load(f)
-            beta = get_stats_steep(betas, mag)
+            temp = get_stats_steep(betas, mag)
             t2s.append(t2)
-            beta_criticals.append(beta)
+            temp_criticals.append(temp)
     t2s = np.array(t2s)
-    beta_criticals = np.array(beta_criticals)
+    temp_criticals = np.array(temp_criticals)
     argsort = np.argsort(t2s)
-    return t2s[argsort], beta_criticals[argsort]
+    return t2s[argsort], temp_criticals[argsort]
 
 def plot(name):
     t2s, transitions = get_phase(f"{name}-")
 
     fig, ax_transitions = plt.subplots(figsize=(6, 4))
 
-    ax_transitions.scatter(t2s, 1/transitions, color="C0")
-    ax_transitions.plot(t2s, 1/transitions, color="C0")
+    ax_transitions.scatter(t2s, transitions, color="C0")
+    ax_transitions.plot(t2s, transitions, color="C0")
     ax_transitions.set_xlabel("t2")
     ax_transitions.set_ylabel("Tc")
 
@@ -40,12 +40,18 @@ betas, mag, sus = load("ising-square.dat")
 print("Square Ising", get_stats_steep(betas, mag))
 betas, mag, sus = load("xy-square.dat")
 print("Square XY", get_stats_steep(betas, mag))
+betas, mag, sus = load("heisenberg-square.dat")
+print("Square Heisenberg", get_stats_steep(betas, mag))
 betas, mag, sus = load("ising-penrose.dat")
 print("Penrose Ising", get_stats_steep(betas, mag))
 betas, mag, sus = load("xy-penrose.dat")
 print("Penrose XY", get_stats_steep(betas, mag))
+betas, mag, sus = load("heisenberg-penrose.dat")
+print("Penrose Heisenberg", get_stats_steep(betas, mag))
 
 plot("rect-ising")
 plot("einstein-ising")
 plot("rect-xy")
 plot("einstein-xy")
+plot("rect-heisenberg")
+plot("einstein-heisenberg")

### 2023-06-09 python/plot.py

@@ -1,57 +1,44 @@
 import matplotlib.pyplot as plt
 import numpy as np
-from util import load
+from util import load, get_stats_steep
+plt.style.use("root")
 
-# plt.rcParams["text.usetex"] = True
-plt.rcParams["font.size"] = 12
-plt.rcParams["font.family"] = "Courier"
-    
-fig, (ax_mag, ax_sus) = plt.subplots(figsize=(6, 7), nrows=2, sharex=True)
-lines = [
-    ("ising-square.dat", "crimson", "dodgerblue", "o"),
-    ("xy-square.dat", "lightsalmon", "cyan", "^"),
-]
-for filename, red, blue, marker in lines:
-    betas, mag, sus = load(filename)
-    ax_mag.plot(1/betas, mag, color=blue, zorder=-10)
-    ax_mag.scatter(1/betas, mag, facecolor=blue, edgecolor='k', marker=marker)
-    ax_sus.plot(1/betas, sus, color=red, zorder=-10)
-    ax_sus.scatter(1/betas, sus, facecolor=red, edgecolor='k', marker=marker)
-ax_sus.set_xlabel("Critical temperature $T_c$")
-ax_mag.set_ylabel("Magnetization $M$")
-ax_sus.set_ylabel("Susceptibility $\chi$")
-ax_sus.set_ylim(1e-1,None)
-ax_mag.set_ylim(0,1)
-ax_sus.set_yscale('log')
-ax_sus.set_xlim(1/betas[-1], 1/betas[0])
-ax_sus.plot([],[],color='k', marker='o', markerfacecolor='lightgray', markeredgecolor='k',label="Ising model")
-ax_sus.plot([],[],color='k', marker='^', markerfacecolor='lightgray', markeredgecolor='k',label="XY model")
-fig.legend(ncol=2, loc="upper center")
-# fig.tight_layout()
-fig.savefig("../figs/square.png", bbox_inches="tight")
+def display_crystal(ising_name, xy_name, save_name):
+    fig, (ax_mag, ax_sus) = plt.subplots(figsize=(6, 7), nrows=2, sharex=True)
+    lines = [
+        (f"{ising_name}.dat", "crimson", "dodgerblue", "o"),
+        (f"{xy_name}.dat", "lightsalmon", "cyan", "^"),
+    ]
+    t_crits = []
+    for filename, red, blue, marker in lines:
+        if filename[:4] == "None": continue
[…]

### 2023-06-09 src/funcs.rs

@@ -19,7 +19,7 @@ pub fn rect_fn(row_size: usize, t2: f32) -> impl Fn(usize) -> SmallVec<(usize, f
         if site % row_size != row_size - 1 {
             neighbors.push((site + 1, 1.))
         } else {
-            neighbors.push((site - row_size + 1, 1.))
+            neighbors.push((site + 1 - row_size, 1.))
         }
         if site / row_size != 0 {
             neighbors.push((site - row_size, t2))

### 2023-06-09 src/lattice.rs

@@ -3,7 +3,7 @@ use std::fmt::Display;
 use crate::spin::{Spin, SmallVec};
 use rand::random;
 
-const CLUSTER_QUEUE_MAX_SIZE: usize = 256_000;
+const CLUSTER_STACK_MAX_SIZE: usize = 256_000;
 const BURN_IN: usize = 128;
 const RUNS_PER_TRIAL: usize = 16;
 
@@ -23,10 +23,11 @@ impl Display for Report {
 pub struct Lattice<S: Spin> {
     data: Vec<S>,
     neighbors: Vec<SmallVec<(usize, f32)>>,
-    cluster_queue: [usize; CLUSTER_QUEUE_MAX_SIZE],
+    cluster_stack: [usize; CLUSTER_STACK_MAX_SIZE],
 }
 
 impl<S: Spin> Lattice<S> {
+    /// Create a new lattice from a function that generates neighbors.
     pub fn new(num_sites: usize, neighbor_func: &impl Fn(usize) -> SmallVec<(usize, f32)>) -> Self {
         let mut data = Vec::with_capacity(num_sites);
         let mut neighbors = Vec::with_capacity(num_sites);
@@ -38,16 +39,18 @@ impl<S: Spin> Lattice<S> {
         Self {
             data,
             neighbors,
-            cluster_queue: [0; CLUSTER_QUEUE_MAX_SIZE],
+            cluster_stack: [0; CLUSTER_STACK_MAX_SIZE],
         }
     }
 
+    /// Zero out the data set
     pub fn zero(&mut self) {
         for item in self.data.iter_mut() {
             *item = S::start();
         }
     }
 
+    /// Run cluster flips and computes the average magnetization and susceptibility every `RUNS_PER_TRIAL` iterations. Returns these as the data set after `n_trials` flips have elapsed, or until `checkpoint.0` flips have elapsed AND magnetization is over `checkpoint.1`. The checkpoint is a runtime-improvement feature. To turn it off, set `checkpoint.1` greater than 1.
     pub fn run(&mut self, beta: f32, n_trials: usize, checkpoint: (usize, f32)) -> Report {
         let mut num_avgs = n_trials - BURN_IN;
 
@@ -76,6 +79,7 @@ impl<S: Spin> Lattice<S> {
         }
     }
 
+    /// Flips exactly one cluster. This code has been optimized to make it branchless. 
     fn evolve(&mut self, beta: f32, count: usize) -> f32 {
[…]

### 2023-06-09 src/main.rs

@@ -1,16 +1,18 @@
 #![allow(dead_code)]
 #![allow(unused_imports)]
 mod lattice;
+mod quantum;
 mod spin;
 mod funcs;
 use core::num;
 use std::{fs::File, io::Write, thread, sync::Arc};
 
 use lattice::Lattice;
+use quantum::QLattice;
 use spin::{Ising, Spin};
 use funcs::{square_fn, load_penrose, load_einstein, rect_fn};
 
-use crate::spin::XY;
+use crate::spin::{XY, Heisenberg};
 
 const NUM_THREADS: usize = 8;
 
@@ -83,6 +85,24 @@ fn one_pass<S: Spin>(lattice: &mut Lattice<S>, start_temp: f32, end_temp: f32,
     }
 }
 
+fn qone_pass(lattice: &mut QLattice, start_temp: f32, end_temp: f32,
+    n_trials: usize, num_betas: usize) -> Data {
+
+    let mut susceptibilities = Vec::new();
+    let mut magnetizations = Vec::new();
+    let betas = reciprocal_linspace(start_temp, end_temp, num_betas);
+    for beta in &betas {
+        println!("Temp {}", 1./beta);
+        lattice.zero();
+        let result = lattice.run(*beta, n_trials, (n_trials / 10, 0.5));
+        magnetizations.push(result.magnetization);
+        susceptibilities.push(result.susceptibility);
+    }
+    Data {
+        betas, magnetizations, susceptibilities
+    }
+}
+
 fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
     layers: usize, n_trials: usize, count_per_iteration: usize) -> Data {
 
@@ -133,18 +153,31 @@ fn main() {
     // ising_rect();
     // println!("Ising einstein");
     // ising_einstein();
+    // println!("heisenberg rect");
+    // heisenberg_rect();
+    // println!("heisenberg einstein");
+    // heisenberg_einstein();
 }
 
 fn one() {
-    let (size, func) = (128*128, square_fn(128));
+    let (size, func) = (12*12, square_fn(12));
     println!("{} sites", size);
-    let mut lattice = Lattice::<Ising>::new(size, &func);
-    let data = one_pass(&mut lattice, 0.01, 4., 1000, 50);
-    data.save("ising-square");
+    let mut lattice = QLattice::new(size, &func);
[…]

### 2023-06-09 src/quantum.rs

@@ -1,0 +1,275 @@
+use crate::spin::{Spin, SmallVec};
+use crate::lattice::Report;
+
+use std::fmt::Display;
+use rand::random;
+
+const CLUSTER_STACK_MAX_SIZE: usize = 256_000;
+const BURN_IN: usize = 4;
+const RUNS_PER_TRIAL: usize = 8;
+const SPIN_BC: bool = true;
+const EPSILON : f32 = 1e-10;
+
+pub struct QLattice {
+    length_three: f32,
+    data: Vec<Vec<f32>>,
+    neighbors: Vec<SmallVec<(usize, f32)>>,
+    cluster_stack: [(usize, f32); CLUSTER_STACK_MAX_SIZE],
+}
+
+fn verify_sorted(l: &Vec<f32>) -> bool {
+    let mut last = None;
+    for i in l {
+        if let Some(las) = last {
+            if *i < las {
+                return false;
+            }
+        }
+        last = Some(*i)
+    }
+    true
+}
+
+impl QLattice {
+    /// Create a new lattice from a function that generates neighbors.
+    pub fn new(num_sites: usize, neighbor_func: &impl Fn(usize) -> SmallVec<(usize, f32)>) -> Self {
+        let mut data = Vec::with_capacity(num_sites);
+        let mut neighbors = Vec::with_capacity(num_sites);
+        let length_three = 2. * (num_sites as f32).sqrt();
+        for site_index in 0..num_sites {
+            data.push(Vec::new());
+            neighbors.push(neighbor_func(site_index));
+        }
+
+        Self {
+            length_three,
+            data,
+            neighbors,
+            cluster_stack: [(0, 0.0); CLUSTER_STACK_MAX_SIZE],
+        }
+    }
+
+    /// Zero out the data set
+    pub fn zero(&mut self) {
+        for item in self.data.iter_mut() {
+            item.clear();
+            // *item = (0..self.length_three as usize).into_iter().map(|_| { random::<f32>() * self.length_three }).collect::<Vec<_>>();
+            item.sort_by(|a,b| a.partial_cmp(b).unwrap() );
+        }
+    }
+
+    /// Same run function as for `Lattice`.
+    pub fn run(&mut self, beta: f32, n_trials: usize, checkpoint: (usize, f32)) -> Report {
+        let mut num_avgs = n_trials - BURN_IN;
+
+        self.evolve(beta, RUNS_PER_TRIAL * BURN_IN);
+        let mut avg_mag = 0.0;
[…]

### 2023-06-09 src/spin.rs

@@ -17,6 +17,13 @@ pub struct XY {
     y: f32,
 }
 
+#[derive(Debug)]
+pub struct Heisenberg {
+    x: f32,
+    y: f32,
+    z: f32,
+}
+
 impl<'a> AddAssign<&'a Self> for Ising {
     fn add_assign(&mut self, rhs: &'a Self) {
         self.data += rhs.data;
@@ -27,6 +34,14 @@ impl<'a> AddAssign<&'a Self> for XY {
     fn add_assign(&mut self, rhs: &'a Self) {
         self.x += rhs.x;
         self.y += rhs.y;
+    }
+}
+
+impl<'a> AddAssign<&'a Self> for Heisenberg {
+    fn add_assign(&mut self, rhs: &'a Self) {
+        self.x += rhs.x;
+        self.y += rhs.y;
+        self.z += rhs.z;
     }
 }
 
@@ -80,6 +95,40 @@ impl Spin for XY {
     }
     fn zero() -> Self {
         Self { x: 0., y: 0. }
+    }
+}
+
+impl Spin for Heisenberg {
+    type V = (f32, f32, f32);
+    fn dot(&mut self, vec: &Self::V) -> f32 {
+        self.x * vec.0 + self.y * vec.1 + self.z * vec.2
+    }
+    fn flip(&mut self, vec: &Self::V, success: bool) {
+        let dot = (vec.0 * self.x + vec.1 * self.y + vec.2 * self.z) * (success as i32 as f32);
+        self.x -= 2. * dot * vec.0;
+        self.y -= 2. * dot * vec.1;
+        self.z -= 2. * dot * vec.2;
+        let norm = 1./(self.x*self.x + self.y*self.y + self.z*self.z).sqrt();
+        self.x *= norm;
+        self.y *= norm;
+        self.z *= norm;
+    }
+    fn norm(&self) -> f32 {
+        (self.x * self.x + self.y * self.y + self.z * self.z).sqrt()
+    }
+    fn random_vec() -> Self::V {
+        let x = 0.1 * thread_rng().sample::<f32,_>(StandardNormal);
+        let y = 0.1 * thread_rng().sample::<f32,_>(StandardNormal);
+        let z = 0.1 * thread_rng().sample::<f32,_>(StandardNormal);
+        let norm = 1./(x*x + y*y + z*z).sqrt();
+        (x * norm, y * norm, z * norm)
+    }
+    fn start() -> Self {
[…]

### 2023-06-09 paper/main.tex
Working on paper

@@
 \begin{abstract}
-  A shape known as the ``einstein tile'' has been discovered which tiles the plane aperiodically.
-We define the Ising, XY, and transverse Ising models on the einstein tiling and compute the magnetization and susceptibility of the tiling as a function of temperature using Monte Carlo methods.
+  A family of shapes known as ``einstein tiles'' has been discovered which tile the plane aperiodically.
+We define the Ising, XY, and Heisenberg models on the einstein tilings and compute the magnetization and susceptibility of the tiling as a function of temperature using Monte Carlo methods.
 This allows us to produce phase diagrams of the relationship between the critical temperature of the tiling and the dimensions of the einstein tile.
@@
 
-\section{Phase Transitions and Spin Models}
+\section{Introduction}
 The Ising model was one of the first statistical mechanics models to contain a phase transition, and remains one of the only exactly solvable examples.
@@
 \end{equation}
-where the brackets indicate an average over the statistical ensemble of states, weighted by the Boltzmann factor $e^{-\beta H}$. Another interesting parameter is the susceptibility
-\begin{equation}
-  \chi = \frac{\partial M}{\partial h} = \frac{1}{N} \braket{\sum_{i} (s_i - M)^2}.
-\end{equation}
-Now that the susceptibility has been defined, we will set $h$ to zero and study the Ising model in the zero magnetic field limit.
+where the brackets indicate an average over the statistical ensemble of states, weighted by the Boltzmann factor $e^{-\beta H}$.
 
-The reason why a phase transition occurs is as follows.
-The Hamiltonian displays $s_i\rightarrow -s_i$ symmetry when when $h=0$.
+The phase transition occurs for the following reason.
+For $h=0$, the Hamiltonian displays $s_i\rightarrow -s_i$ symmetry.
 If the thermodynamic system explores all of phase space, then the average thermodynamic state should possess the same symmetries of the Hamiltonian so that $M=0$.
@@
 
-In addition to the Ising model, we study the XY model which consists of promoting $s_i$ to a two-dimensional unit vector defined at every lattice point.
-The interaction term in the Hamiltonian becomes $\bm s_i \cdot \bm s_j$.
-This XY model has different $T_c$ than the Ising model and can even contain vortices, which are topologically protected quasiparticles where the curl of $\bm s_i$ is nonzero.
-A third spin model is the Transverse Ising Model (TIM), which consists of replacing $s_i$ in Eq.~\ref{eqn:ising} with the Pauli matrices $\bm \sigma_i$ and $\bm \sigma_j$ and interpreting the result as a quantum Hamiltonian.
-Past research \jtd{citation needed} has shown that the $d$-dimensional TIM is equivalent to a classical Ising model in $d+1$ dimensions in the limit of a very strong interaction strength in the extra dimension.
-Thus, critical temperatures for all three of these models can be found using the same techniques.
+The solution to a spin model depends on the lattice on which the model is defined.
+For example, an exact solution exists for the two-dimensional Ising model defined on the square lattice which yields an exact formula for the critical temperature.
+In the absence of a full solution, a series expansions in $\beta$ (valid at high temperature) or $e^{-\beta}$ (valid at low temperature) can be written for the model's thermodynamical quantities.
+For self-dual lattices such as the square lattice, the coefficients of the high- and low-temperature expansions are the same --- a phenomenon known as Kramers-Wannier duality \cite{kramers}.
+This duality leads to an analytical solution for the critical temperature even in the absence of a full solution.
+Even for lattices which are not self-dual, it is sometimes possible to use a similar technique to relate the series expansions of the lattice and its dual to each other.
+This can be done for the hexagonal lattice, for example, which is dual to the triangular lattice \cite{bhattacharjee1995fifty}.
 
-In this final project, we will use Monte Carlo methods discussed in the next section to evaluate the critical temperature of these three models for the recently discovered einstein tiling (figure \ref{fig:einstein}).
-The einstein tile (German for ``one stone'') is the first shape known to tile the plane in a non-repeating pattern \cite{smith2023aperiodic}.
-Such aperiodic tilings are called quasicrystals, and a two-tile version known as the Penrose tiling has gained much traction in the literature \jtd{mention its critical points.
-Also show the einstein tile}.
+In more complicated crystals, however, the only available approaches to computing $T_c$ are numerical.
+For example, the Penrose lattice consists of two tiles which fill the two-dimensional plane aperiodically, making it difficult to model the system analytically.
+Ref.
+\cite{penrose-ising} computes magnetism as a function of temperature using numerical methods and finds that the aperiodic nature of the lattice does not affect the physics at a qualitative level, since $M(T)$ inherits most of its properties from the very general Landau theory.
+
+Recently, the first \textit{single} tile which fills the plane aperiodically has been discovered, and named the einstein tile (German for ``one stone'') (figure \ref{fig:einstein}) \cite{smith2023aperiodic}.
 An interesting fact about the einstein tile is that its dimensions are adjustable.
@@
 If we treat the spin-spin interaction as weaker across the longer bonds, then this $a/b$ ratio represents a free parameter of the crystal, so that the critical temperature is a function of $a/b$.
-Generally, we take the interaction strength $J$ across a bond of length $\ell$ to be $J = J_0 \ell^{-2}$ where $J_0$ is constant.
+The phase diagram of $T_c(a/b)$ is therefore an intrinsic property of the einstein tiling, more so than the commonplace behavior of $M(T)$.
 
@@
 
+In this final project, we use numerical methods to compute the phase diagram $T_c(a/b)$ for the einstein tile under the Ising model and two similar models.
+We also compute the phase diagram for a rectangular lattice with bond lengths $a$ and $b$ to compare to the einstein lattice phase diagram, later using this calculation to solve for the one-dimensional transverse Ising model phase transition.
+The structure of the paper is as follows: in section \ref{sec:spin-models}, we briefly review the spin models.
+In section \ref{sec:wolff} we discuss the numerical methods used to compute $T_c$.
+We present the phase diagrams in section \ref{sec:results}, followed by section \ref{sec:tim} on the transverse Ising model.
+We conclude with section \ref{sec:conclusion}.
+
+\section{Spin models}
+\label{sec:spin-models}
+
+In addition to the Ising model Hamiltonian (Eq.~\ref{eqn:ising}), we study the XY model which consists of promoting $s_i$ to a two-dimensional unit vector defined at every lattice point.
+The interaction term in the Hamiltonian becomes $\bm s_i \cdot \bm s_j$.
+This XY model has different $T_c$ than the Ising model and can even contain vortices, which are topologically protected quasi-particles where the curl of $\bm s_i$ is nonzero.
+A third spin model is the Heisenberg model, which is similar to the XY model except that $s_i$ is promoted to a three-dimensional unit vector.
+Unlike the XY model, the Heisenberg model may not display a true phase transition in the thermodynamic limit \cite{tomita2014finite}.
+Nevertheless, it has phase-transition-like behavior in that $M$ experiences a large drop in a small temperature range.
+The Ising, XY, and Heisenberg models have progressively more degrees of freedom, and their critical temperatures are therefore progressively lower.
+
+A fourth spin model is the Transverse Ising Model (TIM), which consists of replacing $s_i$ in Eq.~\ref{eqn:ising} with the Pauli matrices $\bm \sigma_i$ and $\bm \sigma_j$ and interpreting the result as a quantum Hamiltonian.
+A simple proof using the path integral approach (appendix \ref{app:tim-equivalence}) shows that the $d$-dimensional TIM is equivalent to a classical Ising model in $d+1$ dimensions in the limit of a very strong interaction strength in the extra dimension.
+Thus, critical temperatures for all four of these models can be found using the same techniques.
+
+Beyond the magnetization $M$, another interesting property of a spin model is its susceptibility $\chi = \frac{dM}{dh}$.
+Even in a model which sets $h=0$, $\chi$ an be defined as follows by the fluctuation-dissipation theorem (appendix \ref{app:fluctuation-dissipation}):
+\begin{equation}
+  \chi = \frac{\partial M}{\partial h} = \frac{1}{N}\braket{\sum_i(s_i - M)^2}.
+\end{equation}
+The susceptibility diverges at the critical temperature and otherwise falls off to zero.
+
 \section{Cluster Monte Carlo Methods: the Wolff Algorithm}
-Solving the Ising model is generally difficult.
-Ising himself solved the Ising model for a one-dimensional regular crystal, and a solution was produced 20 years later for a two dimensional square lattice.
-The critical temperature can sometimes be worked out by a clever trick for the Ising model when the lattice is self-dual, such as the square lattice \jtd{cite}.
-Even when the lattice is not self-dual, if the lattice shares enough properties with its dual the connection can be used to determine the critical temperature \jtd{cite}.
+\label{sec:wolff}
 
-However, the einstein lattice (like many lattices) satisfies none of these properties, so that we are reduced to numerically computing the ensemble average of Eq.~\ref{eqn:magnetization}.
+To find the critical temperature of the einstein lattice, we must numerically compute the ensemble average of Eq.~\ref{eqn:magnetization}.
 This average is extremely expensive to compute because the number of states in the ensemble is $2^N$ for the Ising model.
@@
 
-\section{Results}
+\section{Phase Diagrams}
+\label{sec:results}
 
@@
 
-The upper left panel of Figure \ref{fig:square} shows the magnetization $M$ and susceptibility $\chi$ of the Ising, XY, and TI models executed on a square lattice as a function of the lattice width $L = \sqrt{N}$.
-The critical temperature of the Ising model on this lattice is known to be $T_c = \frac{2}{\ln (1 + \sqrt{2})} \approx 2.27$ in units where $k_B = J = 1$ due to the Onsager solution, and indeed the magnetization falls sharpest near this location.
-For the XY model, the critical temperature is approximately $0.8816$, which is likewise close to the drop in $M$.
+The upper left panel of Figure \ref{fig:square} shows the magnetization $M$ and susceptibility $\chi$ of the Ising, XY, and Heisenberg models executed on a square lattice as a function of the lattice width $L = \sqrt{N}$.
+The magnetization drops rapidly near $T_c$, shown as vertical dotted lines.
 The location of these transitions is not perfect, and $M$ it is not strictly zero for $T<T_c$ due to finite-size effects.
 Larger lattice sizes lead to smaller $M$ for $T<T_c$.
-In the XY model the critical temperature is substantially lower due to the increased degrees of freedom of the spin system.
-Furthermore, $M$ approaches its low-temperature limit of 1 linearly for low temperature rather than exponentially, which is a qualitative difference between the Ising and XY models.
+The XY and Heisenberg models have substantially lower $T_c$ due to the increased degrees of freedom of the spin system.
+Furthermore, $M$ approaches its low-temperature limit of 1 linearly for the Heisenberg and XY models, while the approach is exponential for the Ising model.
+This is a qualitative difference between the Ising and XY models.
 
@@
   \centering
-  \includegraphics[width=0.49\linewidth]{../figs/square.png}\hfill
-  \includegraphics[width=0.49\linewidth]{../figs/penrose.png}
+  \includegraphics[width=0.49\linewidth]{../figs/square.pdf}\hfill
+  \includegraphics[width=0.49\linewidth]{../figs/penrose.pdf}
   \caption{Magnetization and susceptibility as a function of $T_c$ under the XY and Ising models.
@@
 For the sake of approximating the critical temperature, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects the $M=1$ axis.
-This approximation results in $T_c\approx 2.26$ for the Ising model on a square lattice and $T_c\approx 0.922$, within a few percent of the true result.
+The critical temperatures obtained from this approximation are shown in table \ref{tab:tc}.
+These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$, and for the XY model MCMC methods give $T_c \approx 0.8816$.
+The Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which cannot display true phase transitions until $N\rightarrow \infty$.
+\begin{table}
+  \centering
+  \begin{tabular}{c|cc}
+    \hline \hline
+    & Square & Penrose \\ \hline
+    Ising & 2.26 & 2.36\\
+    XY & 0.928 & 0.943\\
+    Heisenberg & 0.493 & 0.462\\ \hline \hline
+  \end{tabular}
+  \caption{Critical temperatures for the Ising, XY, and Heisenberg models defined on the square and Penrose lattices. These values are very close to the true values (see text).}
+  \label{tab:tc}
+\end{table}
 
-To demonstrate that the $M$-$T$ relationship is not strongly dependent on the lattice, we also compute the same quantities for the Penrose lattice, shown in the right two panels of figure \ref{fig:square}.
+To demonstrate that the $M(T)$ relationship is not strongly dependent on the lattice, we also compute the same quantities for the Penrose lattice, shown in the right two panels of figure \ref{fig:square}.
 We see the same behavior in how $M$ and $\chi$ approach $T=0$ and $\infty$.
@@
 The major difference between the two is the actual value of $T_c$, which is slightly higher for Penrose.
-\jtd{Why?
-State TC and figure out a way to put errors on it?
-Also draw TC on the plot.}
+One potential explanation is that the average Penrose lattice site is connected to more bonds than the average square lattice site.
+This could lead to a larger effective spin coupling which allows the critical temperature to rive.
 
@@
 Assigning different interaction strengths $J_1$ and $J_2$ to the two bond lengths gives a free parameter $J_1/J_2$.
-Figure \ref{fig:ising-rect} shows the critical temperature as a function of $J_1/J_2$ where we have fixed the tile area to be constant: $J_1J_2 = 1$.
+If we let $J \propto \ell^{-2}$ where $\ell$ is the bond length, then the phase diagram is further parameterized by the length ratio $a/b$.
 
-A similar procedure shows the phase diagram for the einstein tiling, with the ratio of side lengths $a/b$ as the free parameter
+Let us define the anisotropy $\eta$ of the lattice as 
+\begin{equation}
+  \frac{a}{b} = \frac{2}{\pi}\tan^{-1}\eta,\qquad \eta = \tan\frac{\pi a}{2b}.
+\end{equation}
+We fix the product of the lengths $ab=1$.
+Then $\eta \in (-1,1)$ and $\eta=0$ implies $a=b$.
+For the square lattice in particular, rotational symmetry guarantees that the phase diagram $T_c(\eta)$ is symmetric under $a\leftrightarrow b$, or $\eta \rightarrow -\eta$.
+
+Figure \ref{fig:rect} shows the critical temperature as a function of $\eta$ under the Ising, XY, and Heisenberg models. As expected, the result is symmetric around $\eta \rightarrow -\eta$.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{../figs/rect-phase.pdf}
+  \caption{The phase diagram of the rectangular lattice as a function of the anisotropy $\eta$, where $\eta = 0$ is the square lattice and $\eta\rightarrow \pm 1$ is the anisotropic limit.}
+  \label{fig:rect}
+\end{figure}
+
+A similar procedure shows the phase diagram for the einstein tiling, where $a$ and $b$ are now the two lengths shown in figure \ref{fig:einstein}.
+
+\section{Transverse Ising Model}
+\label{sec:tim}
 
 \section{Conclusion}
+\label{sec:conclusion}
+
+\appendix
+
+\section{The Fluctuation-Dissipation Theorem and Magnetic Susceptibility}
+\label{app:fluctuation-dissipation}
+
+\section{Equivalence Between the Transverse Ising Model and Classical Ising Model}
+\label{app:tim-equivalence}
 

### 2023-06-09 python/phase.py

@@ -4,37 +4,37 @@ import os
 from util import load, get_stats_steep
 
 WINDOW = (0.25, 0.55)
+plt.style.use("root")
 
+def post_process(load_result, eta):
+    betas, mag, sus = load_result
+    t2 = np.tan((eta + 1.) / 4. * np.pi)
+    t1 = 1 / t2
+    return (betas / t1, mag, sus)
 
 def get_phase(start_name):
-    t2s = []
+    etas = []
     temp_criticals = []
     for f in os.listdir("../data/output/"):
         if f.startswith(start_name):
-            t2 = float(f[len(start_name):-4])
-            betas, mag, sus = load(f)
+            eta = float(f[len(start_name):-4])
+            betas, mag, sus = post_process(load(f), eta)
             temp = get_stats_steep(betas, mag)
-            t2s.append(t2)
+            etas.append(eta)
             temp_criticals.append(temp)
-    t2s = np.array(t2s)
+    etas = np.array(etas)
     temp_criticals = np.array(temp_criticals)
-    argsort = np.argsort(t2s)
-    return t2s[argsort], temp_criticals[argsort]
+    argsort = np.argsort(etas)
+    return etas[argsort], temp_criticals[argsort]
 
-def plot(name):
-    t2s, transitions = get_phase(f"{name}-")
+def plot(ax, name, label, color, marker):
+    etas, transitions = get_phase(f"{name}-")
 
-    fig, ax_transitions = plt.subplots(figsize=(6, 4))
-
-    ax_transitions.scatter(t2s, transitions, color="C0")
-    ax_transitions.plot(t2s, transitions, color="C0")
-    ax_transitions.set_xlabel("t2")
-    ax_transitions.set_ylabel("Tc")
-
-    # ax_transitions.set_xlim(0.1,2)
-    # ax_transitions.set_ylim(0,1)
-    fig.tight_layout()
-    fig.savefig(f"../figs/{name}-phase.png")
+    ax.plot(etas, transitions, color=color, marker=marker, markerfacecolor=color, markeredgecolor='k', zorder=-1, label=label)
+    ax.set_xlabel("Anisotropy $\eta$")
+    ax.set_ylabel("Critical Temperature $T_c$")
+    ax.set_xlim(-1,1)
+    ax.set_ylim(0,None)
[…]

### 2023-06-09 python/plot.py

@@ -3,11 +3,12 @@ import numpy as np
 from util import load, get_stats_steep
 plt.style.use("root")
 
-def display_crystal(ising_name, xy_name, save_name):
+def display_crystal(ising_name, xy_name, heisenberg_name, save_name):
     fig, (ax_mag, ax_sus) = plt.subplots(figsize=(6, 7), nrows=2, sharex=True)
     lines = [
-        (f"{ising_name}.dat", "crimson", "dodgerblue", "o"),
-        (f"{xy_name}.dat", "lightsalmon", "cyan", "^"),
+        (f"{ising_name}.dat", "maroon", "mediumblue", "o"),
+        (f"{xy_name}.dat", "crimson", "dodgerblue", "^"),
+        (f"{heisenberg_name}.dat", "salmon", "cyan", "d"),
     ]
     t_crits = []
     for filename, red, blue, marker in lines:
@@ -25,20 +26,20 @@ def display_crystal(ising_name, xy_name, save_name):
     ax_mag.set_ylabel("Magnetization $M$")
     ax_sus.set_ylabel("Susceptibility $\chi$")
     ax_sus.set_ylim(1e-1,None)
-    # ax_mag.set_ylim(0,1)
+    ax_mag.set_ylim(0,1)
     ax_sus.set_yscale('log')
     ax_sus.set_xlim(1/betas[-1], 1/betas[0])
     for t_crit, (_,_,_,marker) in zip(t_crits, lines):
-        ax_mag.scatter([t_crit], ax_mag.get_ylim()[1], facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
-        ax_sus.scatter([t_crit], ax_sus.get_ylim()[0], facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
+        ax_mag.scatter([t_crit], ax_mag.get_ylim()[1], s=48,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
+        ax_sus.scatter([t_crit], ax_sus.get_ylim()[0], s=48,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
     ax_sus.plot([],[],color='k', marker='o', markerfacecolor='lightgray',label="Ising model")
     ax_sus.plot([],[],color='k', marker='^', markerfacecolor='lightgray',label="XY model")
-    fig.legend(ncol=2, loc="upper center", bbox_to_anchor=(0.5, 0.95))
[…]

### 2023-06-09 src/funcs.rs

@@ -22,14 +22,14 @@ pub fn rect_fn(row_size: usize, t2: f32) -> impl Fn(usize) -> SmallVec<(usize, f
             neighbors.push((site + 1 - row_size, 1.))
         }
         if site / row_size != 0 {
-            neighbors.push((site - row_size, t2))
+            neighbors.push((site - row_size, 1./(t2 * t2)))
         } else {
-            neighbors.push((row_size * (row_size-1) + site, t2))
+            neighbors.push((row_size * (row_size-1) + site, 1./(t2 * t2)))
         }
         if site / row_size != row_size - 1 {
-            neighbors.push((site + row_size, t2))
+            neighbors.push((site + row_size, 1./(t2 * t2)))
         } else {
-            neighbors.push((site - row_size * (site / row_size), t2))
+            neighbors.push((site - row_size * (site / row_size), 1./(t2 * t2)))
         }
         neighbors
     }
@@ -96,10 +96,10 @@ pub fn load_einstein(name: &str, t2: f32) -> (usize, impl Fn(usize) ->SmallVec<(
             }
             for pair in medium_indices.chunks(2) {
                 if pair[0] as usize == site {
-                    neighbors.push((pair[1] as usize, t2));
+                    neighbors.push((pair[1] as usize, 1./(t2*t2)));
                 }
                 if pair[1] as usize == site {
-                    neighbors.push((pair[0] as usize, t2));
+                    neighbors.push((pair[0] as usize, 1./(t2*t2)));
                 }
             }
             for pair in large_indices.chunks(2) {

### 2023-06-09 src/main.rs

@@ -9,10 +9,8 @@ use std::{fs::File, io::Write, thread, sync::Arc};
 
 use lattice::Lattice;
 use quantum::QLattice;
-use spin::{Ising, Spin};
+use spin::{Ising, XY, Heisenberg, Spin};
 use funcs::{square_fn, load_penrose, load_einstein, rect_fn};
-
-use crate::spin::{XY, Heisenberg};
 
 const NUM_THREADS: usize = 8;
 
@@ -56,6 +54,10 @@ fn reciprocal_linspace(start: f32, end: f32, count: usize) -> Vec<f32> {
     linspace(end, start, count).into_iter().map(|f|{1./f}).collect::<Vec<_>>()
 }
 
+fn get_t2_from_eta(eta: f32) -> f32 {
+    ((eta + 1.) / 4. * std::f32::consts::PI).tan()
+}
+
 fn linspace_ex(start: f32, end: f32, count: usize) -> Vec<f32> {
     if count == 1 {
         return vec![start];
@@ -92,7 +94,7 @@ fn qone_pass(lattice: &mut QLattice, start_temp: f32, end_temp: f32,
     let mut magnetizations = Vec::new();
     let betas = reciprocal_linspace(start_temp, end_temp, num_betas);
     for beta in &betas {
-        println!("Temp {}", 1./beta);
+        print!("Temp {} \t", 1./beta);
         lattice.zero();
         let result = lattice.run(*beta, n_trials, (n_trials / 10, 0.5));
         magnetizations.push(result.magnetization);
@@ -144,29 +146,25 @@ fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
 }
 
 fn main() {
-    one();
-    // println!("XY rect");
-    // xy_rect();
-    // println!("XY einstein");
-    // xy_einstein();
-    // println!("Ising rect");
-    // ising_rect();
-    // println!("Ising einstein");
-    // ising_einstein();
-    // println!("heisenberg rect");
-    // heisenberg_rect();
-    // println!("heisenberg einstein");
-    // heisenberg_einstein();
+    // one();
+    // rect::<Ising>(3.);
+    einstein::<Ising>(0.8);
+    rect::<XY>(1.3);
+    einstein::<XY>(1.5);
+    rect::<Heisenberg>(0.9);
+    einstein::<Heisenberg>(0.4);
 }
 
 fn one() {
     let (size, func) = (12*12, square_fn(12));
[…]

### 2023-06-09 src/quantum.rs

@@ -8,7 +8,7 @@ const CLUSTER_STACK_MAX_SIZE: usize = 256_000;
 const BURN_IN: usize = 4;
 const RUNS_PER_TRIAL: usize = 8;
 const SPIN_BC: bool = true;
-const EPSILON : f32 = 1e-10;
+const EPSILON : f32 = 1e-5;
 
 pub struct QLattice {
     length_three: f32,
@@ -152,21 +152,17 @@ impl QLattice {
                 };
 
                 // 3. Flip spins in this column
-                let mut lr = -random::<f32>().ln() * beta;
-                let mut ll = -random::<f32>().ln() * beta;
+                let mut lr = -random::<f32>().ln();
+                let mut ll = -random::<f32>().ln();
 
-                if stack_pointer > 0 {
-                    for (site, height) in &self.cluster_stack[0..((stack_pointer - 1) as usize)] {
-                        // Do not allow the flipping of points in the queue
-                        if *site != my_index { continue; }
-                        if *height > my_height && *height < my_height + lr {
-                            lr = height - my_height - EPSILON;
-                            println!("Shortened A");
-                        }
-                        if *height < my_height && *height > my_height - ll {
-                            ll = my_height - height - EPSILON;
-                            println!("Shortened B");
-                        }
+                for (site, height) in &self.cluster_stack[0..stack_pointer as usize] {
+                    // Do not allow the flipping of points in the queue
+                    if *site != my_index { continue; }
+                    if *height > my_height && *height < my_height + lr {
+                        lr = height - my_height - EPSILON;
+                    }
+                    if *height < my_height && *height > my_height - ll {
+                        ll = my_height - height - EPSILON;
                     }
                 }
 
@@ -197,24 +193,19 @@ impl QLattice {
                     }
                 }
                 // assert_eq!(verify_sorted(my_column), true);
-                let cr = dl.min(dr);
+                let cr = dr.min(lr);
                 let cl = dl.min(ll);
 
-                println!("{} {} {} {} {} {}", cr + cl, ll, lr, dl, dr, my_height);
-
-                // println!("{} {:?}", my_height, my_column);
-                
                 // 4. Make bridges
                 for (neighbor, strength) in self.neighbors[my_index].iter_mut() {
                     if marked_sites[*neighbor] { continue; }
                     let neighbor_column = &self.data[*neighbor];
                     let mut cluster_done_so_far = 0.;
                     loop {
-                        let lx = -random::<f32>().ln() / (2. * strength);
[…]

### 2023-06-09 src/spin.rs

@@ -65,6 +65,9 @@ impl Spin for Ising {
     fn zero() -> Self {
         Self { data: 0 }
     }
+    fn name() -> String {
+        "ising".to_owned()
+    }
 }
 
 impl Spin for XY {
@@ -95,6 +98,9 @@ impl Spin for XY {
     }
     fn zero() -> Self {
         Self { x: 0., y: 0. }
+    }
+    fn name() -> String {
+        "xy".to_owned()
     }
 }
 
@@ -129,6 +135,9 @@ impl Spin for Heisenberg {
     }
     fn zero() -> Self {
         Self { x: 0., y: 0., z: 0. }
+    }
+    fn name() -> String {
+        "heisenberg".to_owned()
     }
 }
 
@@ -198,4 +207,6 @@ pub trait Spin: for <'a> AddAssign<&'a Self> + Sized + Debug {
     fn start() -> Self;
     /// Get the norm for the sake of averaging
     fn norm(&self) -> f32;
+    /// Get the name of this spin model
+    fn name() -> String;
 }

### 2023-06-11 paper/main.tex
Worked on paper

@@
 \usepackage{graphicx}
+\usepackage{braket}
 \usepackage[colorlinks=true]{hyperref}
 
-\newcommand{\braket}[1]{\left \langle #1 \right \rangle}
 \newcommand{\parens}[1]{\left ( #1 \right )}
+\newcommand{\brackets}[1]{\left [ #1 \right ]}
 \newcommand{\jtd}[1]{{\color{red}\textbf{JTD: #1}}}
@@
 \begin{equation}
-  H = -J \sum_{\braket{ij}} s_i s_j + h \sum_i s_i
+  H = -J \sum_{\braket{ij}} s_i s_j - h \sum_i s_i
   \label{eqn:ising}
@@
 A fourth spin model is the Transverse Ising Model (TIM), which consists of replacing $s_i$ in Eq.~\ref{eqn:ising} with the Pauli matrices $\bm \sigma_i$ and $\bm \sigma_j$ and interpreting the result as a quantum Hamiltonian.
-A simple proof using the path integral approach (appendix \ref{app:tim-equivalence}) shows that the $d$-dimensional TIM is equivalent to a classical Ising model in $d+1$ dimensions in the limit of a very strong interaction strength in the extra dimension.
+A simple proof using the path integral approach (section \ref{sec:tim}) shows that the $d$-dimensional TIM is equivalent to a classical Ising model in $d+1$ dimensions in the limit of a very strong interaction strength in the extra dimension.
 Thus, critical temperatures for all four of these models can be found using the same techniques.
@@
 Beyond the magnetization $M$, another interesting property of a spin model is its susceptibility $\chi = \frac{dM}{dh}$.
-Even in a model which sets $h=0$, $\chi$ an be defined as follows by the fluctuation-dissipation theorem (appendix \ref{app:fluctuation-dissipation}):
+Even in a model which sets $h=0$, $\chi$ can be achieved by explicitly differentiating Eq.~\ref{eqn:magnetization}, including the $e^{-\beta H}$ weight in the average.
+This differentiation results in the zero-$h$ limit of
 \begin{equation}
@@
 \end{equation}
-The susceptibility diverges at the critical temperature and otherwise falls off to zero.
+This is a special case of the fluctuation-dissipation theorem, which states that a driving field $h$ causes fluctuations in the driven variable $M$ which are proportional to the susceptibility $\chi$.
+A spin model is most susceptible to magnetic field near $T_c$, because the presence of a magnetic field changes the value of $T_c$ and therefore can greatly change the magnetization of a system near $T_c$.
+In particular, the susceptibility diverges at the critical temperature and otherwise falls off to zero.
 
@@
 
-We implemented the Wolff algorithm with a custom code base \footnote{\jtd{LINK}} and executed for a variety of spin models and crystals.
+We implemented the Wolff algorithm with a custom code base \footnote{\url{https://github.com/jack-dinsmore/quasing-model}} and executed for a variety of spin models and crystals.
 Except when stated otherwise, we used a finite-size lattice with initial conditions of random spins.
@@
 In both crystals and both models, the point of maximum susceptibility and maximum $dM/dT$ occur at approximately the same location.
-For the sake of approximating the critical temperature, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects the $M=1$ axis.
+For the sake of approximating the critical temperature, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects $M=0.5$.
 The critical temperatures obtained from this approximation are shown in table \ref{tab:tc}.
-These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$, and for the XY model MCMC methods give $T_c \approx 0.8816$.
+These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$ in units where $J=k_B=1$, and for the XY model MCMC methods give $T_c \approx 0.8816$.
 The Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which cannot display true phase transitions until $N\rightarrow \infty$.
@@
   \end{tabular}
-  \caption{Critical temperatures for the Ising, XY, and Heisenberg models defined on the square and Penrose lattices. These values are very close to the true values (see text).}
+  \caption{Critical temperatures for the Ising, XY, and Heisenberg models defined on the square and Penrose lattices in units where $J = k_B = 1$.
+These values are very close to the true values (see text).}
   \label{tab:tc}
@@
 \begin{equation}
-  \frac{a}{b} = \frac{2}{\pi}\tan^{-1}\eta,\qquad \eta = \tan\frac{\pi a}{2b}.
+  \frac{a}{b} = \tan\parens{\pi\frac{\eta + 1}4},\qquad \eta = \frac{4}{\pi}\tan^{-1}\parens{\frac{a}{b}}-1.
+  \label{eqn:anisotropy}
 \end{equation}
@@
 
-Figure \ref{fig:rect} shows the critical temperature as a function of $\eta$ under the Ising, XY, and Heisenberg models. As expected, the result is symmetric around $\eta \rightarrow -\eta$.
+Figure \ref{fig:rect-phase} shows the critical temperature as a function of $\eta$ under the Ising, XY, and Heisenberg models. As expected, the result is symmetric around $\eta \rightarrow -\eta$.
 
@@
   \includegraphics[width=\linewidth]{../figs/rect-phase.pdf}
-  \caption{The phase diagram of the rectangular lattice as a function of the anisotropy $\eta$, where $\eta = 0$ is the square lattice and $\eta\rightarrow \pm 1$ is the anisotropic limit.}
-  \label{fig:rect}
+  \caption{The phase diagram of the rectangular lattice as a function of the anisotropy $\eta$, where $\eta = 0$ is the square lattice and $\eta\rightarrow \pm 1$ is the anisotropic limit.
+This function is even due to the rotational symmetry of a rectangular lattice.}
+  \label{fig:rect-phase}
 \end{figure}
@@
 A similar procedure shows the phase diagram for the einstein tiling, where $a$ and $b$ are now the two lengths shown in figure \ref{fig:einstein}.
+The phase diagram is shown in figure \ref{fig:einstein-phase}.
+
+\jtd{Discussion}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{../figs/einstein-phase.pdf}
+  \caption{The phase diagram of the einstein tiling as a function of the anisotropy $\eta$. The einstein lattice depicted in figure \ref{fig:einstein} had $\eta = 1/\sqrt{3}\approx 0.577$.}
+  \label{fig:einstein-phase}
+\end{figure}
 
@@
 
+Quantum mechanical models such as the TIM cannot in general be represented by the same formalism as a classical model.
+However the the TIM is a special case: A $d$-dimensional TIM is equivalent to a $d+1$-dimensional Ising model.
+We will prove this in the case of $d=1$.
+To see this, consider the quantum partition function
+\begin{equation}
+  Z = \mathrm{Tr}\parens{e^{-\tau H}}.
+  \label{eqn:quantum-partition}
+\end{equation}
+For clarity, the Hamiltonian for the TIM is defined as  
+\begin{equation}
+  H = -J \sum_{\braket{ij}}\bm \sigma_i \cdot \bm \sigma_j - h \sum_i\sigma_i^3
+\end{equation}
+where $\bm \sigma_i$ is the spin operator (a vector of Pauli matrices) for site $i$. Since $H$ is time-independent and therefore self-commuting, this is 
+\begin{equation}
+  \begin{split}
+    Z =& \mathrm{Tr}\parens{e^{-\epsilon H}\dots e^{-\epsilon H}}\\
+     =& \sum_{\psi_0\in\mathcal{H}}\bra{\psi_0}e^{-\epsilon H}\parens{\sum_{\psi_1\in\mathcal{H}}\ket{\psi_1}\bra{\psi_1}}\times\dots\\
+     &\times e^{-i\epsilon H}\parens{\sum_{\psi_n\in\mathcal{H}}\ket{\psi_n}\bra{\psi_n}}\ket{\psi_0}
+  \end{split}
+\end{equation}
+where $\ket{\psi_a}$ indicates a quantum state in the Hilbert space $\mathcal{H}$.
+In the second line we just inserted the identity many times as well as the definition of the trace.
+The number $n$ is set such that $\epsilon n = \beta$ to satisfy Eq.~\ref{eqn:quantum-partition}.
+Setting $\epsilon$ sufficiently low, we can expand the exponential to first order in $\epsilon$:
+\begin{equation}
+  \begin{split}
+    Z =& \sum_{\{\psi\}}\bra{\psi_0}(1 - \epsilon H)\ket{\psi_1}\bra{\psi_1}\times\dots\\
+    &\times (1 - \epsilon H)\ket{\psi_n}\braket{\psi_n|\psi_0}.
+  \end{split}
+  \label{eqn:partition-simplified}
+\end{equation}
+where we have collapsed all the sums into one.
+In an orthonormal basis where $\braket{\psi_a|\psi_b} = \delta_{ab}$, the only $\mathcal{O}(\epsilon^0)$ term in Eq.~\ref{eqn:partition-simplified} is for all $\ket{\psi_a}$ equal.
+So to order $\mathcal{O}(\epsilon)$,
+\begin{equation}
+  \begin{split}
+    Z &= 1 - \epsilon\sum_{a\neq b}\braket{\psi_a|H|\psi_b}\\
+    Z &= 1 - \epsilon\sum_{a\neq b}\Braket{\psi_a|-J\sum_{\braket{ij}}\bm \sigma_i \cdot \bm \sigma_j - h\sum_i\sigma_i^3|\psi_b}
+  \end{split}
+\end{equation}
+As our basis, we may take the wavefunctions $\ket{s_1 s_2 \dots s_N}$ where $s_i$ indicates the sign of the spin of site $i$ and there are $N$ sites.
+Explicit calculation verifies that $\braket{s_is_j|\bm \sigma_i \cdot \bm \sigma_j|s_i's_j'}=\delta_{s_i,s_j}\delta_{s'_i,s'_j}$.
+We sum over the $s'$ spins and replace the $s$ Kronecker delta with $s_i s_j$, which is equal to $\delta_{s_i,s_j}$ shifting by a constant.
+Likewise, $\braket{s|\sigma^3|s'}=\delta_{s,s'}\sim ss'$.
+Expressing $\ket{\psi_a}=\ket{s_1 s_2 \dots s_N}$ and $\ket{\psi_b}=\ket{s'_1 s'_2 \dots s'_N}$, we get
+\begin{equation}
+  Z = 1 - \epsilon\sum_{\{s\},\{s'\}} \brackets{-J\sum_{\braket{ij}}\bm s_is_j - h\sum_is_i s'_i}.
+\end{equation}
+Since this expression is order $\epsilon$, we may write it as an exponential:
+\begin{equation}
+  \begin{split}
+    Z &= \sum_{\{s\},\{s'\}} e^{-K}.\\
+    K &= -\epsilon J\sum_{\braket{ij}}\bm s_is_j +\frac{1}{2}\ln(\epsilon h)\sum_is_i s'_i.
+  \end{split}
+\end{equation}
+
+If we stack the $s$ and $s'$ spins into a square grid, then this is the classical partition function of a square Ising model, with different vertical and horizontal interaction strengths.
+The anisotropy is very high: $\eta \sim 1 - C\sqrt{\epsilon}$ for a constant $C$ under the definition of Eq.~\ref{eqn:anisotropy}.
+
+\jtd{Discussion}
+
 \section{Conclusion}
@@
 
-\appendix
-
-\section{The Fluctuation-Dissipation Theorem and Magnetic Susceptibility}
-\label{app:fluctuation-dissipation}
-
-\section{Equivalence Between the Transverse Ising Model and Classical Ising Model}
-\label{app:tim-equivalence}
+\jtd{Conclusion}
 

### 2023-06-11 python/phase.py

@@ -6,19 +6,13 @@ from util import load, get_stats_steep
 WINDOW = (0.25, 0.55)
 plt.style.use("root")
 
-def post_process(load_result, eta):
-    betas, mag, sus = load_result
-    t2 = np.tan((eta + 1.) / 4. * np.pi)
-    t1 = 1 / t2
-    return (betas / t1, mag, sus)
-
 def get_phase(start_name):
     etas = []
     temp_criticals = []
     for f in os.listdir("../data/output/"):
         if f.startswith(start_name):
             eta = float(f[len(start_name):-4])
-            betas, mag, sus = post_process(load(f), eta)
+            betas, mag, sus = load(f)
             temp = get_stats_steep(betas, mag)
             etas.append(eta)
             temp_criticals.append(temp)
@@ -50,16 +44,17 @@ betas, mag, sus = load("heisenberg-penrose.dat")
 print("Penrose Heisenberg", get_stats_steep(betas, mag))
 
 
-fig, ax = plt.subplots(figsize=(6, 4))
+fig, ax = plt.subplots()
 plot(ax, "rect-ising", "Ising", "gold", "o")
 plot(ax, "rect-xy", "XY", "orange", "^")
 plot(ax, "rect-heisenberg", "Heisenberg", "darkgoldenrod", "d")
+ax.set_ylim(0,4)
 fig.legend(ncol=3, loc="upper center", bbox_to_anchor=(0.5, 1.06))
 fig.tight_layout()
 fig.savefig(f"../figs/rect-phase.png")
 fig.savefig(f"../figs/rect-phase.pdf")
 
-fig, ax = plt.subplots(figsize=(6, 4))
+fig, ax = plt.subplots()
 plot(ax, "einstein-ising", "Ising", "gold", "o")
 plot(ax, "einstein-xy", "XY", "orange", "^")
 plot(ax, "einstein-heisenberg", "Heisenberg", "darkgoldenrod", "d")

### 2023-06-11 python/plot.py

@@ -4,7 +4,7 @@ from util import load, get_stats_steep
 plt.style.use("root")
 
 def display_crystal(ising_name, xy_name, heisenberg_name, save_name):
-    fig, (ax_mag, ax_sus) = plt.subplots(figsize=(6, 7), nrows=2, sharex=True)
+    fig, (ax_mag, ax_sus) = plt.subplots(figsize=(9, 11), nrows=2, sharex=True)
     lines = [
         (f"{ising_name}.dat", "maroon", "mediumblue", "o"),
         (f"{xy_name}.dat", "crimson", "dodgerblue", "^"),
@@ -22,7 +22,7 @@ def display_crystal(ising_name, xy_name, heisenberg_name, save_name):
         ax_mag.scatter(1/betas, mag, facecolor=blue, marker=marker)
         ax_sus.plot(1/betas, sus, color=red, zorder=-1)
         ax_sus.scatter(1/betas, sus, facecolor=red, marker=marker)
-    ax_sus.set_xlabel("Critical temperature $T_c$")
+    ax_sus.set_xlabel("Temperature $T$")
     ax_mag.set_ylabel("Magnetization $M$")
     ax_sus.set_ylabel("Susceptibility $\chi$")
     ax_sus.set_ylim(1e-1,None)
@@ -30,8 +30,8 @@ def display_crystal(ising_name, xy_name, heisenberg_name, save_name):
     ax_sus.set_yscale('log')
     ax_sus.set_xlim(1/betas[-1], 1/betas[0])
     for t_crit, (_,_,_,marker) in zip(t_crits, lines):
-        ax_mag.scatter([t_crit], ax_mag.get_ylim()[1], s=48,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
-        ax_sus.scatter([t_crit], ax_sus.get_ylim()[0], s=48,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
+        ax_mag.scatter([t_crit], ax_mag.get_ylim()[1], s=96,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
+        ax_sus.scatter([t_crit], ax_sus.get_ylim()[0], s=96,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
     ax_sus.plot([],[],color='k', marker='o', markerfacecolor='lightgray',label="Ising model")
[…]

### 2023-06-11 src/funcs.rs

@@ -5,21 +5,21 @@ use npy::NpyData;
 use crate::spin::SmallVec;
 
 pub fn square_fn(row_size: usize) -> impl Fn(usize) -> SmallVec<(usize, f32)> {
-    rect_fn(row_size, 1.)
+    rect_fn(row_size, 1., 1.)
 }
 
-pub fn rect_fn(row_size: usize, t2: f32) -> impl Fn(usize) -> SmallVec<(usize, f32)> {
+pub fn rect_fn(row_size: usize, t1: f32, t2: f32) -> impl Fn(usize) -> SmallVec<(usize, f32)> {
     move |site: usize| {
         let mut neighbors = SmallVec::new();
         if site % row_size != 0 {
-            neighbors.push((site - 1, 1.))
+            neighbors.push((site - 1, 1./(t1*t1)))
         } else {
-            neighbors.push((site + row_size - 1, 1.))
+            neighbors.push((site + row_size - 1, 1./(t1*t1)))
         }
         if site % row_size != row_size - 1 {
-            neighbors.push((site + 1, 1.))
+            neighbors.push((site + 1, 1./(t1*t1)))
         } else {
-            neighbors.push((site + 1 - row_size, 1.))
+            neighbors.push((site + 1 - row_size, 1./(t1*t1)))
         }
         if site / row_size != 0 {
             neighbors.push((site - row_size, 1./(t2 * t2)))
@@ -61,7 +61,7 @@ pub fn load_penrose(level: usize) -> (usize, impl Fn(usize) ->SmallVec<(usize, f
     )
 }
 
-pub fn load_einstein(name: &str, t2: f32) -> (usize, impl Fn(usize) ->SmallVec<(usize, f32)>) {
+pub fn load_einstein(name: &str, t1: f32, t2: f32) -> (usize, impl Fn(usize) ->SmallVec<(usize, f32)>) {
     let mut small_buf = vec![];
     let mut med_buf = vec![];
     let mut long_buf = vec![];
@@ -88,10 +88,10 @@ pub fn load_einstein(name: &str, t2: f32) -> (usize, impl Fn(usize) ->SmallVec<(
             let mut neighbors = SmallVec::new();
             for pair in small_indices.chunks(2) {
                 if pair[0] as usize == site {
-                    neighbors.push((pair[1] as usize, 1.0));
+                    neighbors.push((pair[1] as usize, 1.0/(t1*t1)));
                 }
                 if pair[1] as usize == site {
-                    neighbors.push((pair[0] as usize, 1.0));
[…]

### 2023-06-11 src/lattice.rs

@@ -75,7 +75,7 @@ impl<S: Spin> Lattice<S> {
         let sus = (avg_mag2 - avg_mag*avg_mag) * self.data.len() as f32;
         Report {
             magnetization: avg_mag,
-            susceptibility: sus,
+            susceptibility: sus * beta,
         }
     }
 

### 2023-06-11 src/main.rs

@@ -54,8 +54,12 @@ fn reciprocal_linspace(start: f32, end: f32, count: usize) -> Vec<f32> {
     linspace(end, start, count).into_iter().map(|f|{1./f}).collect::<Vec<_>>()
 }
 
+fn get_t1_from_eta(eta: f32) -> f32 {
+    ((-eta + 1.) / 4. * std::f32::consts::PI).tan().powi(2)
+}
+
 fn get_t2_from_eta(eta: f32) -> f32 {
-    ((eta + 1.) / 4. * std::f32::consts::PI).tan()
+    ((eta + 1.) / 4. * std::f32::consts::PI).tan().powi(2)
 }
 
 fn linspace_ex(start: f32, end: f32, count: usize) -> Vec<f32> {
@@ -147,7 +151,7 @@ fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
 
 fn main() {
     // one();
-    // rect::<Ising>(3.);
+    rect::<Ising>(3.);
     einstein::<Ising>(0.8);
     rect::<XY>(1.3);
     einstein::<XY>(1.5);
@@ -156,12 +160,12 @@ fn main() {
 }
 
 fn one() {
-    let (size, func) = (12*12, square_fn(12));
-    println!("{} sites", size);
-    let mut lattice = QLattice::new(size, &func);
-    // let data = qone_pass(&mut lattice, 0.01, 0.01, 1000, 1);
-    let data = qone_pass(&mut lattice, 0.1, 2., 1000, 10);
-    data.save("tim-square");
+    // let (size, func) = (12*12, square_fn(12));
+    // println!("{} sites", size);
+    // let mut lattice = QLattice::new(size, &func);
+    // // let data = qone_pass(&mut lattice, 0.01, 0.01, 1000, 1);
+    // let data = qone_pass(&mut lattice, 0.1, 2., 1000, 10);
+    // data.save("tim-square");
     
     // let (size, func) = (128*128, square_fn(128));
     // println!("{} sites", size);
@@ -177,19 +181,19 @@ fn one() {
     // let data = one_pass(&mut lattice, 0.01, 4., 10000, 50);
     // data.save("heisenberg-square");
 
-    // let (size, func) = load_penrose(9);
-    // println!("{} sites", size);
-    // let mut lattice = Lattice::<Ising>::new(size, &func);
-    // let data = one_pass(&mut lattice, 0.01, 4., 10000, 50);
[…]

### 2023-06-11 src/quantum.rs

@@ -91,7 +91,7 @@ impl QLattice {
         }
         Report {
             magnetization: avg_mag,
-            susceptibility: sus,
+            susceptibility: sus * beta,
         }
     }
 

### 2023-06-11 paper/main.tex
Worked more on paper. Almost done

@@
-\documentclass[amsmath,amssymb,aps,twocolumn,nofootinbib]{revtex4-2}
+\documentclass[
+  amsmath,
+  amssymb,
+  aps,
+  twocolumn,
+  nofootinbib,
+  floatfix,
+]{revtex4-2}
 
@@
 \usepackage{braket}
+\usepackage{xcolor}
 \usepackage[colorlinks=true]{hyperref}
@@
 \begin{equation}
-  M = \frac{1}{N}\braket{\sum_{i} s_i}
+  M = \frac{1}{N}\Braket{\sum_{i} s_i}
   \label{eqn:magnetization}
@@
   \includegraphics[width=\linewidth]{../figs/einstein.pdf}
-  \caption{The einstein tiling}
+  \caption{The einstein tiling with $b/a=\sqrt{3}$. Any other fraction $b/a$ gives a different einstein tiling}
   \label{fig:einstein}
@@
 \begin{equation}
-  \chi = \frac{\partial M}{\partial h} = \frac{1}{N}\braket{\sum_i(s_i - M)^2}.
+  \chi = \frac{\partial M}{\partial h} = \frac{1}{N}\Braket{\sum_i(s_i - M)^2}.
 \end{equation}
@@
 \begin{equation}
-  \frac{a}{b} = \tan\parens{\pi\frac{\eta + 1}4},\qquad \eta = \frac{4}{\pi}\tan^{-1}\parens{\frac{a}{b}}-1.
+  \frac{a}{b} = \tan\brackets{\frac{\pi}{4}\parens{1+\eta}},\qquad \eta = \frac{4}{\pi}\tan^{-1}\parens{\frac{a}{b}}-1.
   \label{eqn:anisotropy}
 \end{equation}
-We fix the product of the lengths $ab=1$.
+We fix the product of the lengths $ab=1$, so that 
+\begin{equation}
+  a^2 = \tan\brackets{\frac{\pi}{4}\parens{1+\eta}},\qquad b^2 = \tan\brackets{\frac{\pi}{4}\parens{1-\eta}}
+\end{equation}
+
+.
 Then $\eta \in (-1,1)$ and $\eta=0$ implies $a=b$.
 For the square lattice in particular, rotational symmetry guarantees that the phase diagram $T_c(\eta)$ is symmetric under $a\leftrightarrow b$, or $\eta \rightarrow -\eta$.
-
-Figure \ref{fig:rect-phase} shows the critical temperature as a function of $\eta$ under the Ising, XY, and Heisenberg models. As expected, the result is symmetric around $\eta \rightarrow -\eta$.
 
@@
   \includegraphics[width=\linewidth]{../figs/rect-phase.pdf}
-  \caption{The phase diagram of the rectangular lattice as a function of the anisotropy $\eta$, where $\eta = 0$ is the square lattice and $\eta\rightarrow \pm 1$ is the anisotropic limit.
+  \caption{The phase diagram of the rectangular lattice as a function of the anisotropy $\eta$, where $\eta = 0$ is the square lattice.
+The Ising and XY model $T_c$ peak at $\eta \approx 1/3$ and $1/4$ respectively, while the Heisenberg model $T_c$ is constant until $\eta \approx 1/6$.
 This function is even due to the rotational symmetry of a rectangular lattice.}
@@
 
+Figure \ref{fig:rect-phase} shows the critical temperature as a function of $\eta$ under the Ising, XY, and Heisenberg models.
+As expected, the result is symmetric around $\eta \rightarrow -\eta$.
+Anisotropy appears to increase the critical temperature near $\eta = 0$ in the Ising and XY models, though this trend must turn over at some $\eta$ to achieve a $T_c\rightarrow 0$ value in the limit as $\eta \rightarrow \pm 1$.
+For the Ising model, the turnover occurs at $\eta=\pm\frac{1}{3}$.
+For the XY model, it occurs at $\eta = \pm \frac{1}{3}$.
+For the Heisenberg model, $T_c$ is constant to within a fraction of a percent until $\eta \approx 0.15$ when $T_c$ begins to decrease.
+All of these turnover values of $\eta$ may depend on the length $L$ of the lattice.
+In the $N\rightarrow \infty$ limit, a series expansion approach may help to reveal why certain values of $\eta$ change the behavior of $T_c$.
+
 A similar procedure shows the phase diagram for the einstein tiling, where $a$ and $b$ are now the two lengths shown in figure \ref{fig:einstein}.
 The phase diagram is shown in figure \ref{fig:einstein-phase}.
+In this case, anisotropy decreases the critical temperature in every model so that the maximum anisotropy occurs at $\eta = 0$, which corresponds to a regular lattice.
+The critical temperature falls so quickly that $\eta = -1/3$, which is the value of $\eta$ depicted in figure \ref{fig:einstein}, already has $T_c \approx 0$.
+This is due to the relatively low coordination number of each site in the einstein tiling; many sites are connected to the minimum two bonds whereas the Penrose and square model had average coordination number of 4 in the $N\rightarrow \infty$ limit.
+Since more connected lattices have stronger spin interactions, this is qualitatively similar to fixing $J$.
+Dimensional analysis requires that the critical value of any crystal is proportional to $\beta_c J = J/T_c$ so that increasing $J$ implies increasing $T_c$.
 
-\jtd{Discussion}
+Surprisingly, $T_c(\eta)$ possesses approximate $\eta\rightarrow -\eta$ symmetry, even though $\eta$ and $-\eta$ represent different tilings.
+This suggests the presence of some greater symmetry in the einstein tiling which has not yet been remarked.
+Other models defined on the einstein tiling do not contain this symmetry.
+The Ising and Heisenberg display slight deviations from $\eta\rightarrow -\eta$ symmetry; it is possible that the symmetry only appears in the XY case.
 
@@
   \includegraphics[width=\linewidth]{../figs/einstein-phase.pdf}
-  \caption{The phase diagram of the einstein tiling as a function of the anisotropy $\eta$. The einstein lattice depicted in figure \ref{fig:einstein} had $\eta = 1/\sqrt{3}\approx 0.577$.}
+  \caption{The phase diagram of the einstein tiling as a function of the anisotropy $\eta$.
+An unexpected $\eta \rightarrow -\eta$ symmetry has appeared.
+The einstein lattice depicted in figure \ref{fig:einstein} had $\eta = -1/3$.}
   \label{fig:einstein-phase}
@@
 Quantum mechanical models such as the TIM cannot in general be represented by the same formalism as a classical model.
-However the the TIM is a special case: A $d$-dimensional TIM is equivalent to a $d+1$-dimensional Ising model.
+However the the TIM is a special case: A $d$-dimensional TIM is equivalent to a $d+1$-dimensional Ising model, as first discovered in Ref.~\cite{schultz64two}.
 We will prove this in the case of $d=1$.
@@
 
-\jtd{Discussion}
+We have already computed the behavior of the two-dimensional classical Ising model; according to figure \ref{fig:rect-phase}, $T_c\rightarrow 0$ as $\eta \rightarrow 1$, so that the critical temperature of the TIM is zero.
+This is a qualitative difference between the TIM and the classical Ising model; in the classial model, the only one of the two minima of the free energy must be occupied, forcing the system to choose either $M=+1$ or $M=-1$.
+In the quantum model both can be simultaneously occupied, with zero-energy fermion-like domain walls separating the two.
+Thus, the ordered phase has magnetization $M=0$ which cannot occur in the classical case.
 
@@
 \label{sec:conclusion}
+After an introduction to phase transitions in classical spin-like models, first in the case of the Ising model and then in the general case of a Landau field theory, we solve the Ising, XY, and Heisenberg models on a rectangular (regular) lattice and the einstein (aperiodic) tiling.
+As our method, we used the Wolff algorithm with moderate resolution and a modest number of iterations.
+Our results are therefore accurate only to a few figures of precision.
 
-\jtd{Conclusion}
+The magnetization and susceptibility as a function of temperature were discussed and shown to be dramatically different between the models, though similar for different crystals.
+This was explained as a consequence of the Landau field theory explanation for phase transitions, which predicts that the behavior of a theory is sensitive to the number of degrees of freedom of the spin field and less to the structure of the lattice.
+
+Phase diagrams of the critical temperature as a function of the anisotropy of the lattice structure were then shown; these are not predicted by Landau field theory and therefore are much more dependent on the lattice structure.
+Anisotropy was seen to increase the critical temperature of the square lattice, but decrease the critical temperature of the einstein tiling.
+The square lattice showed model-dependent turnover values of anisotropy where $T_c$ was stationary.
+The einstein tiling displayed approximate $\eta \rightarrow -\eta$ symmetry despite the fact that $\eta$ and $-\eta$ represent different lattices, indicative of a deeper self-symmetry to the lattice.
+
+Finally, we introduced a quantum model --- the Transverse Ising model (TIM) --- and showed that the one-dimensional TIM it is equivalent to the two-dimensional classical Ising model in the anisotropic limit.
+Our calculations had revealed a critical temperature of zero for this limit, indicating that the TIM is not magnetized at zero temperature.
+We explained that this is because the quantum nature of the TIM allows it to simultaneously occupy both the spin-up and spin-down minima of the free energy in the low-temperature limit.
+Thus, the thermodynamic state is ordered, full of fermionic domain walls which leave an overall magnetization of zero.
 

### 2023-06-11 python/phase.py

@@ -48,7 +48,7 @@ fig, ax = plt.subplots()
 plot(ax, "rect-ising", "Ising", "gold", "o")
 plot(ax, "rect-xy", "XY", "orange", "^")
 plot(ax, "rect-heisenberg", "Heisenberg", "darkgoldenrod", "d")
-ax.set_ylim(0,4)
+# ax.set_ylim(0,6)
 fig.legend(ncol=3, loc="upper center", bbox_to_anchor=(0.5, 1.06))
 fig.tight_layout()
 fig.savefig(f"../figs/rect-phase.png")

### 2023-06-11 python/plot.py

@@ -15,6 +15,7 @@ def display_crystal(ising_name, xy_name, heisenberg_name, save_name):
         if filename[:4] == "None": continue
         betas, mag, sus = load(filename)
         t_crit = get_stats_steep(betas, mag)
+        print(filename, t_crit)
         t_crits.append(t_crit)
         ax_sus.axvline(t_crit, linestyle="dotted", linewidth=1, color='k', zorder=-2)
         ax_mag.axvline(t_crit, linestyle="dotted", linewidth=1, color='k', zorder=-2)
@@ -32,14 +33,15 @@ def display_crystal(ising_name, xy_name, heisenberg_name, save_name):
     for t_crit, (_,_,_,marker) in zip(t_crits, lines):
         ax_mag.scatter([t_crit], ax_mag.get_ylim()[1], s=96,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
         ax_sus.scatter([t_crit], ax_sus.get_ylim()[0], s=96,facecolor="lightgray", clip_on=False, zorder=10, marker=marker)
-    ax_sus.plot([],[],color='k', marker='o', markerfacecolor='lightgray',label="Ising model")
-    ax_sus.plot([],[],color='k', marker='^', markerfacecolor='lightgray',label="XY model")
-    ax_sus.plot([],[],color='k', marker='d', markerfacecolor='lightgray',label="Heisenberg model")
+    ax_sus.plot([],[],color='k', marker='o', markerfacecolor='lightgray',label="Ising")
+    ax_sus.plot([],[],color='k', marker='^', markerfacecolor='lightgray',label="XY")
+    ax_sus.plot([],[],color='k', marker='d', markerfacecolor='lightgray',label="Heisenberg")
     fig.legend(ncol=3, loc="upper center", bbox_to_anchor=(0.5, 0.95))
     plt.subplots_adjust(hspace=0.05)
     fig.savefig(f"../figs/{save_name}.png", bbox_inches="tight")
     fig.savefig(f"../figs/{save_name}.pdf", bbox_inches="tight")
 
-display_crystal("tim-square", None, None, "test")
+display_crystal("tim-square", None, None, "tim")
+display_crystal("rect-ising-0.33000004", None, None, "test")
 display_crystal("ising-square", "xy-square", "heisenberg-square", "square")
 display_crystal("ising-penrose", "xy-penrose", "heisenberg-square", "penrose")

### 2023-06-11 python/util.py

@@ -16,7 +16,7 @@ def load(filename):
     
 def get_stats_steep(betas, mag):
     max_slope = None
-    max_slope_beta = None
+    max_slope_temp = None
     max_slope_mag= None
     temps = 1/betas
     if len(mag) <= 1: return np.nan, 1

### 2023-06-11 src/main.rs

@@ -55,11 +55,11 @@ fn reciprocal_linspace(start: f32, end: f32, count: usize) -> Vec<f32> {
 }
 
 fn get_t1_from_eta(eta: f32) -> f32 {
-    ((-eta + 1.) / 4. * std::f32::consts::PI).tan().powi(2)
+    ((-eta + 1.) / 4. * std::f32::consts::PI).tan()
 }
 
 fn get_t2_from_eta(eta: f32) -> f32 {
-    ((eta + 1.) / 4. * std::f32::consts::PI).tan().powi(2)
+    ((eta + 1.) / 4. * std::f32::consts::PI).tan()
 }
 
 fn linspace_ex(start: f32, end: f32, count: usize) -> Vec<f32> {
@@ -151,7 +151,7 @@ fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
 
 fn main() {
     // one();
-    rect::<Ising>(3.);
+    rect::<Ising>(6.);
     einstein::<Ising>(0.8);
     rect::<XY>(1.3);
     einstein::<XY>(1.5);

### 2023-06-12 paper/main.tex
Finished draft of paper

--- 2023-06-11 paper/main.tex
@@
-\documentclass[amsmath,amssymb,aps,twocolumn,nofootinbib]{revtex4-2}
+\documentclass[
+  amsmath,
+  amssymb,
+  aps,
+  twocolumn,
+  nofootinbib,
+  nolongbibliography,
+  floatfix,
+]{revtex4-2}
 
@@
 \usepackage{braket}
-\usepackage[colorlinks=true]{hyperref}
+\usepackage{xcolor}
+\usepackage[colorlinks=true,allcolors=magenta]{hyperref}
 
@@
 The Ising model was one of the first statistical mechanics models to contain a phase transition, and remains one of the only exactly solvable examples.
-It consists of defining a variable $s_i = \pm 1$ at every site of a lattice with $N$ sites as a classical approximation of the up-or-down spin of a fermion.
-On further analogy with fermionic spins, it asserts that the spins interact by defining the classical Hamiltonian
+It consists of defining a variable $s_i = \pm 1$ at every site of a lattice as a classical approximation of the up-or-down spin of a fermion.
+On further analogy with fermionic spins, it allows neighboring spins to interact under the classical Hamiltonian
 \begin{equation}
@@
 \end{equation}
-where $h$ is an applied magnetic field. This model's phase transition is most visible in the magnetization, defined as
+where $h$ is an applied magnetic field and $\braket{ij}$ indicate neighbors. This model's phase transition is visible in the magnetization, defined as
 \begin{equation}
-  M = \frac{1}{N}\braket{\sum_{i} s_i}
+  M = \frac{1}{N}\Braket{\sum_{i} s_i}
   \label{eqn:magnetization}
 \end{equation}
-where the brackets indicate an average over the statistical ensemble of states, weighted by the Boltzmann factor $e^{-\beta H}$.
+where the brackets indicate a statistical average over all states, weighted by the Boltzmann factor $e^{-\beta H}$.
 
@@
 If the thermodynamic system explores all of phase space, then the average thermodynamic state should possess the same symmetries of the Hamiltonian so that $M=0$.
-However, for low temperature we also expect the system to attain the lowest possible energy.
-This state consists of all spins up (or all spins down) has $M = 1$ (or $-1$), which violates the expectation of no magnetization.
+However, at low temperature we also expect the system to attain the lowest possible energy, which consists of all spins up (or all spins down).
+Such a state has $M =\pm 1$ which violates the expectation of no magnetization.
 The resolution to this apparent contradiction that there exists a critical temperature $T_c$ below which energetic considerations dominate and the crystal is magnetic ($|M|>0$), and above which the symmetry of phase space dominates and the crystal is not magnetic ($M = 0$).
@@
 This approximation is called Landau Field Theory and it yields a form which is readily able to create phase transitions.
-Taking the Ising model as an example, we could approximate the $s_i$ spins defined on discrete points $x_i$ as the field $s(\bm x)$ defined for any position $\bm x$.
+Taking the Ising model as an example, we approximate the $s_i$ spins defined on discrete points $x_i$ as the field $s(\bm x)$ defined for any position $\bm x$.
 The original Hamiltonian contained translational, rotational, and $s\rightarrow-s$ symmetry, and therefore the free energy should as well.
-Expanding it in powers of $s(\bm x)$ guarantees the form
+Expanding its integrand in powers of $s(\bm x)$ guarantees the form
 \begin{equation}
-F \propto -\int d^d x \, \parens{(\nabla s(\bm x))^2 + A s(\bm x)^2 + B s(\bm x)^4 + \dots}
+F \propto -\int d^2 x \brackets{(\nabla s(\bm x))^2 + A s(\bm x)^2 + B s(\bm x)^4 + \dots}
 \label{eqn:landau}
 \end{equation}
-where $d$ is the number of dimensions of the lattice and $A$ and $B$ are arbitrary constants.
-If $A>0$, then free energy is minimized at every point by setting $s(\bm x)$.
-However, if $A<0$, then minimizing the free energy necessitates setting $s(\bm x)$ to some nonzero value, thereby magnetizing the crystal and creating a phase transition.
-In principle, all the parameters of the model including $A$ may depend on temperature.
-All that is necessary to describe a phase transition is therefore to write a form of $A$ which allows it to cross from positive to negative at a critical temperature $T_c$.
+where $A$ and $B$ are arbitrary constants.
+If $A>0$, then free energy is minimized at every point by setting $s(\bm x)=0$.
+However if $A<0$, then the average value of $s(\bm x)0$ is nonzero and the crystal is magnetized.
+In principle, all the parameters of the model including $A$ may depend on temperature, so any theory whose $A$ crosses from positive to negative contains a phase transition.
 
-The utility of this field theory approximation is in its generality.
-All Ising-like models like it are Landau field theories and are all described by Eq~\ref{eqn:landau}, meaning they have the same near-critical-point behavior.
+The utility of this field theory framework is in its generality.
+All Ising-like models are Landau field theories and are all described by Eq~\ref{eqn:landau}, meaning they have the same near-critical-point behavior.
 This statement is independent of the lattice or interaction strengths; it depends only on the field $\bm s$ and the dimensionality of the lattice.
@@
 
-The solution to a spin model depends on the lattice on which the model is defined.
+% The solution to a spin model depends on the lattice on which the model is defined.
 For example, an exact solution exists for the two-dimensional Ising model defined on the square lattice which yields an exact formula for the critical temperature.
@@
 
-In more complicated crystals, however, the only available approaches to computing $T_c$ are numerical.
+In complicated crystals, the only available approaches to computing this $T_c$ are numerical.
 For example, the Penrose lattice consists of two tiles which fill the two-dimensional plane aperiodically, making it difficult to model the system analytically.
 Ref.
-\cite{penrose-ising} computes magnetism as a function of temperature using numerical methods and finds that the aperiodic nature of the lattice does not affect the physics at a qualitative level, since $M(T)$ inherits most of its properties from the very general Landau theory.
+\cite{penrose-ising} computes magnetism as a function of temperature using numerical methods and finds that the aperiodic nature of the lattice does not affect the physics at a qualitative level, since $M(T)$ inherits most of its properties from the general Landau theory.
 
@@
 The tile consists of two side lengths $a$ and $b$, and any ratio of $a/b$ leads to an aperiodic tiling except $a/b=0$, 1, or $\infty$.
-If we treat the spin-spin interaction as weaker across the longer bonds, then this $a/b$ ratio represents a free parameter of the crystal, so that the critical temperature is a function of $a/b$.
-The phase diagram of $T_c(a/b)$ is therefore an intrinsic property of the einstein tiling, more so than the commonplace behavior of $M(T)$.
+If we treat the spin-spin interaction as dependent on bond length, then this $a/b$ ratio represents a free parameter of the model, so that the critical temperature is a function of $a/b$.
+The phase diagram of $T_c(a/b)$ is therefore an intrinsic property of the einstein tiling, and more flexible than the behavior of $M(T)$.
 
@@
 In this final project, we use numerical methods to compute the phase diagram $T_c(a/b)$ for the einstein tile under the Ising model and two similar models.
-We also compute the phase diagram for a rectangular lattice with bond lengths $a$ and $b$ to compare to the einstein lattice phase diagram, later using this calculation to solve for the one-dimensional transverse Ising model phase transition.
-The structure of the paper is as follows: in section \ref{sec:spin-models}, we briefly review the spin models.
-In section \ref{sec:wolff} we discuss the numerical methods used to compute $T_c$.
-We present the phase diagrams in section \ref{sec:results}, followed by section \ref{sec:tim} on the transverse Ising model.
+The spin models are discussed in section \ref{sec:spin-models} and the numerical methods in section \ref{sec:wolff}.
+We also compute the phase diagram for a rectangular lattice to compare to the einstein lattice phase diagram in section \ref{sec:results}, later using this calculation to solve for the one-dimensional transverse Ising model in section \ref{sec:tim}.
 We conclude with section \ref{sec:conclusion}.
@@
 The interaction term in the Hamiltonian becomes $\bm s_i \cdot \bm s_j$.
-This XY model has different $T_c$ than the Ising model and can even contain vortices, which are topologically protected quasi-particles where the curl of $\bm s_i$ is nonzero.
 A third spin model is the Heisenberg model, which is similar to the XY model except that $s_i$ is promoted to a three-dimensional unit vector.
-Unlike the XY model, the Heisenberg model may not display a true phase transition in the thermodynamic limit \cite{tomita2014finite}.
-Nevertheless, it has phase-transition-like behavior in that $M$ experiences a large drop in a small temperature range.
-The Ising, XY, and Heisenberg models have progressively more degrees of freedom, and their critical temperatures are therefore progressively lower.
+These models are all different in that they have progressively more spin degrees of freedom (DOFs), leading to progressively lower $T_c$.
+Each model also has specialized features, like the XY model's topologically protected vortices and the Heisenberg model's pseudo-phase transition \cite{tomita2014finite}, but these are beyond the scope of this report.
 
 A fourth spin model is the Transverse Ising Model (TIM), which consists of replacing $s_i$ in Eq.~\ref{eqn:ising} with the Pauli matrices $\bm \sigma_i$ and $\bm \sigma_j$ and interpreting the result as a quantum Hamiltonian.
-A simple proof using the path integral approach (section \ref{sec:tim}) shows that the $d$-dimensional TIM is equivalent to a classical Ising model in $d+1$ dimensions in the limit of a very strong interaction strength in the extra dimension.
-Thus, critical temperatures for all four of these models can be found using the same techniques.
+A simple proof (section \ref{sec:tim}) shows that the TIM is equivalent to a classical Ising model, so critical temperatures for all four of these models can be found using the same techniques.
 
@@
 A spin model is most susceptible to magnetic field near $T_c$, because the presence of a magnetic field changes the value of $T_c$ and therefore can greatly change the magnetization of a system near $T_c$.
-In particular, the susceptibility diverges at the critical temperature and otherwise falls off to zero.
+In particular, the susceptibility diverges at $T_c$ and otherwise falls off to zero.
 
@@
 Many states are not representative, making it difficult to find a new state $\psi$ which has high enough $P(\psi)$ to be added to $K$.
-One way to resolve this problem is to make $\psi$ very similar to a state $\phi$ already in $K$.
+One way to resolve this problem is to make $\psi$ very similar to a state $\phi$ already in $K$ --- a technique known as a Markov Chain Monte Carlo method (MCMC).
 Assuming $\phi$ is representative of the ensemble, then $\psi$ is also likely to be represented by the ensemble.
-We then add $\psi$ to $K$ with some probability $P(\phi \rightarrow \psi)$, which is a function we discuss in the next paragraph.
-The technique of using previous states $\phi$ to create a new state $\psi$ is called a Markov Chain Monte Carlo method (MCMC).
+We then add $\psi$ to $K$ with some probability $P(\phi \rightarrow \psi)$.
 
@@
 This condition is necessary to prevent $K$ from getting stuck in an unrepresentative region of phase space.
-The probability to go from $\phi$ to $\psi$ is the probability for the MCMC to be at state $\phi$ in the first place times the probability to switch to $\psi$: $P(\phi \rightarrow \psi)P(\phi)$.
+The probability to go from $\phi$ to $\psi$ is the probability for the MCMC to be at state $\phi$ in the first place times the probability to switch to $\psi$.
 Thus, detailed balance requires
@@
 
-For the XY model, Eq.~\ref{eqn:wolff-algo} must be modified to handle vector-valued spins. The approach is to choose a random vector $\bm r$ as a direction against which to compare to spins, so that
+For the XY and Heisenberg models, Eq.~\ref{eqn:wolff-algo} must be modified to handle vector-valued spins.
+The approach is to choose a random unit vector $\bm r$ as a direction against which to compare to spins, so that
 \begin{equation}
@@
 
-A third and final extension is necessary to the Wolff algorithm in order to handle quantum spin models.
+% A third and final extension is necessary to the Wolff algorithm in order to handle quantum spin models.
 As mentioned above, the TIM in two dimensions is equivalent to the Ising model in three dimensions where interaction strengths $J$ are very strong in the new direction.
@@
 
-\section{Phase Diagrams}
+\section{Results}
 \label{sec:results}
@@
 
-The upper left panel of Figure \ref{fig:square} shows the magnetization $M$ and susceptibility $\chi$ of the Ising, XY, and Heisenberg models executed on a square lattice as a function of the lattice width $L = \sqrt{N}$.
-The magnetization drops rapidly near $T_c$, shown as vertical dotted lines.
-The location of these transitions is not perfect, and $M$ it is not strictly zero for $T<T_c$ due to finite-size effects.
-Larger lattice sizes lead to smaller $M$ for $T<T_c$.
-The XY and Heisenberg models have substantially lower $T_c$ due to the increased degrees of freedom of the spin system.
+The upper left panel of Figure \ref{fig:square} shows the magnetization $M$ and susceptibility $\chi$ of the Ising, XY, and Heisenberg models executed on a square lattice with periodic boundary conditions as a function of the temperature $T$.
+The magnetization drops rapidly near $T_c$, shown as vertical dotted lines, though the drop is not immediate due to finite-size effects.
+The XY and Heisenberg models have substantially lower $T_c$ due to the increased DOFs of the spin system.
 Furthermore, $M$ approaches its low-temperature limit of 1 linearly for the Heisenberg and XY models, while the approach is exponential for the Ising model.
@@
 \textit{Left}: A square lattice with $N=128^2=16,384$ sites.
-\textit{Right}: A Penrose lattice with $N=21,316$ sites.}
+\textit{Right}: A Penrose lattice with $N=21,316$ sites.
+The three spin models yield very different predictions but the two crystals are similar due to their field theory explanation.}
   \label{fig:square}
@@
 
-The susceptibility in the bottom left panel produces a maximum at $T_c$, decaying to zero more rapidly on the low-temperature side than high-temperature.
-For an infinite lattice, the peak will be likewise infinite, but it is blunted by finite $N$.
-The peak in susceptibility can be used as a measure of the location of $T_c$, as can the point of maximum slope of $M$.
+The susceptibility $\chi$ in the bottom left panel produces a maximum at $T_c$.
+For the Ising model, the decay to zero more is more rapid on the low-temperature side than high-temperature, though the reverse is true for the XY and Heisenberg models.
+For an infinite lattice, the peak in $\chi$ will be likewise infinite, though here it is blunted by finite $N$.
 
-In both crystals and both models, the point of maximum susceptibility and maximum $dM/dT$ occur at approximately the same location.
-For the sake of approximating the critical temperature, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects $M=0.5$.
+For the sake of approximating $T_c$, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects $M=0.5$.
 The critical temperatures obtained from this approximation are shown in table \ref{tab:tc}.
 These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$ in units where $J=k_B=1$, and for the XY model MCMC methods give $T_c \approx 0.8816$.
-The Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which cannot display true phase transitions until $N\rightarrow \infty$.
+The Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which never have true phase transitions in any case.
 \begin{table}
   \centering
-  \begin{tabular}{c|cc}
+  \begin{tabular}{c|ccc}
     \hline \hline
-    & Square & Penrose \\ \hline
-    Ising & 2.26 & 2.36\\
-    XY & 0.928 & 0.943\\
-    Heisenberg & 0.493 & 0.462\\ \hline \hline
+    & Square & Penrose & Einstein\\ \hline
+    Ising & 2.26 & 2.36 & 0.396\\
+    XY & 0.928 & 0.943 & 0.130\\
+    Heisenberg & 0.493 & 0.462 & 0.0548\\ \hline \hline
   \end{tabular}
-  \caption{Critical temperatures for the Ising, XY, and Heisenberg models defined on the square and Penrose lattices in units where $J = k_B = 1$.
-These values are very close to the true values (see text).}
+  \caption{Critical temperatures $T_c$ for the Ising, XY, and Heisenberg models defined on the square and Penrose lattices in units where $J = k_B = 1$.
+These values are very close to the true values (see text).
+Also shown are $T_c$ for the einstein lattice with $\eta=-\frac{1}{3}$ as depicted in figure \ref{fig:einstein}.}
   \label{tab:tc}
@@
 To demonstrate that the $M(T)$ relationship is not strongly dependent on the lattice, we also compute the same quantities for the Penrose lattice, shown in the right two panels of figure \ref{fig:square}.
-We see the same behavior in how $M$ and $\chi$ approach $T=0$ and $\infty$.
-The behavior near $T_c$ is slightly different in the steepness of the curves due to the difference in the number of sites.
-The major difference between the two is the actual value of $T_c$, which is slightly higher for Penrose.
-One potential explanation is that the average Penrose lattice site is connected to more bonds than the average square lattice site.
-This could lead to a larger effective spin coupling which allows the critical temperature to rive.
+The same limiting behavior of $\chi$ and $M$ is present in both crystals, with slight changes in the steepness of the curves due to the difference in $N$.
+The major difference between the two is the actual value of $T_c$, which is slightly higher for Penrose (table \ref{tab:tc}).
+This confirms the result of Ref.~\cite{penrose-ising}.
 
@@
 \begin{equation}
-  \frac{a}{b} = \tan\parens{\pi\frac{\eta + 1}4},\qquad \eta = \frac{4}{\pi}\tan^{-1}\parens{\frac{a}{b}}-1.
+  \frac{a}{b} = \tan\brackets{\frac{\pi}{4}\parens{1+\eta}},\qquad \eta = \frac{4}{\pi}\tan^{-1}\parens{\frac{a}{b}}-1.
   \label{eqn:anisotropy}
 \end{equation}
-We fix the product of the lengths $ab=1$.
+We fix the product of the lengths $ab=1$, so that 
+\begin{equation}
+  a^2 = \tan\brackets{\frac{\pi}{4}\parens{1+\eta}},\qquad b^2 = \tan\brackets{\frac{\pi}{4}\parens{1-\eta}}.
+\end{equation}
 Then $\eta \in (-1,1)$ and $\eta=0$ implies $a=b$.
 For the square lattice in particular, rotational symmetry guarantees that the phase diagram $T_c(\eta)$ is symmetric under $a\leftrightarrow b$, or $\eta \rightarrow -\eta$.
-
-Figure \ref{fig:rect-phase} shows the critical temperature as a function of $\eta$ under the Ising, XY, and Heisenberg models. As expected, the result is symmetric around $\eta \rightarrow -\eta$.
 
@@
   \includegraphics[width=\linewidth]{../figs/rect-phase.pdf}
-  \caption{The phase diagram of the rectangular lattice as a function of the anisotropy $\eta$, where $\eta = 0$ is the square lattice and $\eta\rightarrow \pm 1$ is the anisotropic limit.
+  \caption{The phase diagram of the rectangular lattice as a function of the anisotropy $\eta$, where $\eta = 0$ is the square lattice.
 This function is even due to the rotational symmetry of a rectangular lattice.}
@@
 
-A similar procedure shows the phase diagram for the einstein tiling, where $a$ and $b$ are now the two lengths shown in figure \ref{fig:einstein}.
-The phase diagram is shown in figure \ref{fig:einstein-phase}.
-
-\jtd{Discussion}
+Figure \ref{fig:rect-phase} shows $T_c$ as a function of $\eta$ for the rectangular lattice under the Ising, XY, and Heisenberg models.
+As expected, the result is symmetric around $\eta \rightarrow -\eta$.
+Anisotropy appears to increase $T_c$ near $\eta = 0$ in the Ising and XY models, though this trend turns over at some $\eta$ to achieve the $T_c\rightarrow 0$ value in the limit as $\eta \rightarrow \pm 1$.
+For the Ising model, the turnover occurs at $\eta\approx \pm 0.625$ after which $T_c$ falls to zero rapidly.
+For the XY model, the maximum value of $T_c$ occurs at $\eta \approx \pm 0.362$, which is half the turnover $\eta$ of the Ising model.
+However, $T_c$ falls only slowly until $\eta \approx 0.5$, after which $T_c$ drops to zero.
 
@@
   \includegraphics[width=\linewidth]{../figs/einstein-phase.pdf}
-  \caption{The phase diagram of the einstein tiling as a function of the anisotropy $\eta$. The einstein lattice depicted in figure \ref{fig:einstein} had $\eta = 1/\sqrt{3}\approx 0.577$.}
+  \caption{The phase diagram of the einstein tiling as a function of the anisotropy $\eta$. The temperature is even after a shift in $\eta$.}
   \label{fig:einstein-phase}
 \end{figure}
+
+For the Heisenberg model, $T_c$ is maximal at $\eta=0$, though it remains constant to within a fraction of a percent until $\eta \approx 1/3$ at which point $T_c$ begins to decrease.
+All of these turnover values of $\eta$ may depend on the length $L$ of the lattice.
+In the $N\rightarrow \infty$ limit, a series expansion approach may help to reveal why these turnover values occur and how they depend on the number of DOFs of the spin model.
+
+A similar procedure shows the phase diagram for the einstein tiling with $N=7,047$ sites, where $a$ and $b$ are now the two lengths shown in figure \ref{fig:einstein}.
+The phase diagram is shown in figure \ref{fig:einstein-phase}.
+In this case, anisotropy decreases $T_c$ in every model so that the maximum anisotropy occurs near $\eta = 0$, which corresponds to a regular lattice.
+The $T_c$ of the lattice depicted in figure \ref{fig:einstein}, which has $\eta = -1/3$, is given in table \ref{tab:tc}.
+Critical temperatures for all $\eta$ are much lower than for any other crystals studied here because of the relatively low coordination number of each site in the einstein tiling.
+Many sites are connected to the minimum two bonds in einstein whereas the Penrose and square model had average coordination number of 4 in the $N\rightarrow \infty$ limit.
+Since more connected lattices have stronger spin interactions, this is qualitatively similar to fixing $J$.
+Dimensional analysis requires that the critical value of any crystal is proportional to $\beta_c J = J/T_c$ so that increasing $J$ implies increasing $T_c$.
+
+If $\eta$ is shifted by some amount $\eta \rightarrow \eta - x$, then $T_c(\eta)$ possesses $\eta\rightarrow -\eta$ symmetry.
+This is surprising because $\eta$ and $-\eta$ correspond to completely different lattices, suggesting the presence of some greater symmetry in the einstein tiling which has not yet been remarked.
+Other models defined on the einstein tiling do not contain this symmetry.
+The shift value $x$ is 0.029 for the Ising model, 0.052 for the XY model, and 0.033 for the Heisenberg model, though these shift values may depend on both numerical fluctuations of the MCMC and on the finite $N$ used.
 
@@
 \end{equation}
-For clarity, the Hamiltonian for the TIM is defined as  
+For clarity, the Hamiltonian for the TIM is
 \begin{equation}
@@
 If we stack the $s$ and $s'$ spins into a square grid, then this is the classical partition function of a square Ising model, with different vertical and horizontal interaction strengths.
-The anisotropy is very high: $\eta \sim 1 - C\sqrt{\epsilon}$ for a constant $C$ under the definition of Eq.~\ref{eqn:anisotropy}.
+The anisotropy is very high: $\eta \sim 1 - C\sqrt{\epsilon}$ for some constant $C$.
 
-\jtd{Discussion}
+We have already computed the behavior of the two-dimensional classical Ising model; according to figure \ref{fig:rect-phase}, $T_c\rightarrow 0$ as $\eta \rightarrow 1$, so that the critical temperature of the TIM is zero.
+This is a qualitative difference between the TIM and the classical Ising model; in the classial model, the only one of the two minima of the free energy must be occupied, forcing the system to choose either $M=+1$ or $M=-1$.
+In the quantum model both can be simultaneously occupied, with zero-energy fermion-like domain walls separating the two.
+Thus, the ordered phase has magnetization $M=0$ which cannot occur in the classical case.
 
@@
 \label{sec:conclusion}
+After an introduction to phase transitions in classical spin-like models, first in the case of the Ising model and then in the general case of a Landau field theory, we solve the Ising, XY, and Heisenberg models on a rectangular lattice and the aperiodic Penrose and einstein tilings.
 
-\jtd{Conclusion}
+The magnetization and susceptibility as a function of temperature were discussed and shown to be dramatically different between the models, though similar for different crystals.
+This was explained as a consequence of the Landau field theory explanation for phase transitions, which predicts that the behavior of a theory is sensitive to the number of degrees of freedom of the spin field and less to the structure of the lattice.
+
+Phase diagrams of the critical temperature as a function of the lattice anisotropy were shown; these are not predicted by Landau field theory and are therefore less theoretically constrained.
+Anisotropy was seen to increase the critical temperature of the square lattice, but decrease the critical temperature of the einstein tiling.
+The rectangular lattice achieved maximum critical temperature at model-dependent ``turnover values'' of anisotropy, and the einstein tiling displayed unexpected symmetry after a model-dependent shift in the anisotropy.
+Understanding either of these features may require deeper investigation.
+
+Finally, we introduced a quantum Ising model --- the Transverse Ising model (TIM) --- and showed that the one-dimensional TIM it is equivalent to a two-dimensional classical anisotropic Ising model.
+We used our two-dimensional Ising model calculations to discuss the properties of the TIM and qualitatively contrasted the TIM with the classical models investigated here.
 

### 2023-06-12 python/phase.py

@@ -6,6 +6,25 @@ from util import load, get_stats_steep
 WINDOW = (0.25, 0.55)
 plt.style.use("root")
 
+def middle(a,b,c):
+    # Returns the center of the quadratic interpolation between three points
+    num = a[0]**2 * (b[1] - c[1]) + b[0]**2 * (c[1] - a[1]) + c[0]**2 * (a[1] - b[1])
+    denom = a[1] * (b[0] - c[0]) + b[1] * (c[0] - a[0]) + c[1] * (a[0] - b[0])
+    return -0.5 * num / denom
+
+def max_middle(xs, ys):
+    s = np.array([xs,ys]).transpose()
+    argmax = np.argmax(ys)
+    return middle(s[argmax-1], s[argmax], s[argmax+1])
+    
+def max_left_right(xs, ys):
+    res = 0
+    for mask in [xs > 0, xs < 0]:
+        argmax = np.argmax(ys[mask])
+        s = np.array([xs,ys]).transpose()[mask]
+        res += np.abs(middle(s[argmax-1], s[argmax], s[argmax+1]))
+    return res / 2
+
 def get_phase(start_name):
     etas = []
     temp_criticals = []
@@ -13,7 +32,7 @@ def get_phase(start_name):
         if f.startswith(start_name):
             eta = float(f[len(start_name):-4])
             betas, mag, sus = load(f)
-            temp = get_stats_steep(betas, mag)
+            temp = max(0,get_stats_steep(betas, mag))
             etas.append(eta)
             temp_criticals.append(temp)
     etas = np.array(etas)
@@ -23,6 +42,12 @@ def get_phase(start_name):
 
 def plot(ax, name, label, color, marker):
     etas, transitions = get_phase(f"{name}-")
+
+    if name in ["rect-ising", "rect-xy"]:
+        print(name, max_left_right(etas, transitions))
+
+    if name in ["einstein-ising", "einstein-xy", "einstein-heisenberg"]:
+        print(name, max_middle(etas, transitions))
 
     ax.plot(etas, transitions, color=color, marker=marker, markerfacecolor=color, markeredgecolor='k', zorder=-1, label=label)
     ax.set_xlabel("Anisotropy $\eta$")
@@ -42,13 +67,20 @@ betas, mag, sus = load("xy-penrose.dat")
 print("Penrose XY", get_stats_steep(betas, mag))
[…]

### 2023-06-12 src/main.rs

@@ -151,12 +151,12 @@ fn search<S: Spin>(lattice: &mut Lattice<S>, bottom: f32, top: f32,
 
 fn main() {
     // one();
-    rect::<Ising>(6.);
+    // rect::<Ising>(6.);
     einstein::<Ising>(0.8);
-    rect::<XY>(1.3);
-    einstein::<XY>(1.5);
-    rect::<Heisenberg>(0.9);
-    einstein::<Heisenberg>(0.4);
+    // rect::<XY>(1.3);
+    // einstein::<XY>(1.5);
+    // rect::<Heisenberg>(0.9);
+    // einstein::<Heisenberg>(0.4);
 }
 
 fn one() {
@@ -243,6 +243,7 @@ fn einstein<S: Spin>(tmax: f32) {
             for eta in eta_chunk {
                 println!("{}", eta);
                 let (size, func) = load_einstein("7k", get_t1_from_eta(eta), get_t2_from_eta(eta));
+                println!("{}", size);
                 let mut lattice = Lattice::<S>::new(size, &func);
                 let data = one_pass(&mut lattice, 0.01, tmax, 10000, 50);
                 data.save(&format!("einstein-{}-{:.8}", S::name(), eta));

### 2023-06-12 paper/main.tex
Done

@@
 \section{Introduction}
-The Ising model was one of the first statistical mechanics models to contain a phase transition, and remains one of the only exactly solvable examples.
+The Ising model was one of the first statistical mechanics models to contain a phase transition and remains one of the only exactly solvable examples.
 It consists of defining a variable $s_i = \pm 1$ at every site of a lattice as a classical approximation of the up-or-down spin of a fermion.
@@
 \end{equation}
-where $h$ is an applied magnetic field and $\braket{ij}$ indicate neighbors. This model's phase transition is visible in the magnetization, defined as
+where $h$ is an applied magnetic field and $\braket{ij}$ indicate neighbors. Henceforth, we'll set $h=0$. This model's phase transition is visible in the magnetization, defined as
 \begin{equation}
@@
 The phenomenon of critical temperature has been generalized to other theories.
-In particular, if the system's variables are arranged as a members of a field $s(\bm x)$, then the free energy may be approximated as an integral over functions of $s(\bm x)$.
+In particular, if the system's variables are arranged as a members of a field $s(\bm x)$, then the free energy may be approximated as an integral of a function of $s(\bm x)$.
 This approximation is called Landau Field Theory and it yields a form which is readily able to create phase transitions.
@@
 where $A$ and $B$ are arbitrary constants.
-If $A>0$, then free energy is minimized at every point by setting $s(\bm x)=0$.
-However if $A<0$, then the average value of $s(\bm x)0$ is nonzero and the crystal is magnetized.
-In principle, all the parameters of the model including $A$ may depend on temperature, so any theory whose $A$ crosses from positive to negative contains a phase transition.
+If $A>0$, then the integrand is minimized by setting $s(\bm x)=0$.
+However if $A<0$, then the average value of $s(\bm x)$ is nonzero and the crystal is magnetized.
+In principle, all the parameters of the model including $A$ may depend on temperature, so any theory whose $A$ crosses from positive to negative at a temperature $T_c$ contains a phase transition.
 
-The utility of this field theory framework is in its generality.
+The utility of this field theory framework is its generality.
 All Ising-like models are Landau field theories and are all described by Eq~\ref{eqn:landau}, meaning they have the same near-critical-point behavior.
-This statement is independent of the lattice or interaction strengths; it depends only on the field $\bm s$ and the dimensionality of the lattice.
+This statement is independent of the lattice or interaction strengths; it depends only on the degrees of freedom (DOFs) of the field $s$ and the dimensionality of the lattice.
 A particular lattice's only unique properties its its values of $A$ and $B$; i.e., the value of its critical temperature.
@@
 
-In complicated crystals, the only available approaches to computing this $T_c$ are numerical.
+In complicated crystals, the only available approaches to computing $T_c$ are numerical.
 For example, the Penrose lattice consists of two tiles which fill the two-dimensional plane aperiodically, making it difficult to model the system analytically.
@@
 The spin models are discussed in section \ref{sec:spin-models} and the numerical methods in section \ref{sec:wolff}.
-We also compute the phase diagram for a rectangular lattice to compare to the einstein lattice phase diagram in section \ref{sec:results}, later using this calculation to solve for the one-dimensional transverse Ising model in section \ref{sec:tim}.
+We also compute the phase diagram for a rectangular lattice to compare to the einstein lattice phase diagram in section \ref{sec:results}, later using this calculation to explain the one-dimensional transverse Ising model in section \ref{sec:tim}.
 We conclude with section \ref{sec:conclusion}.
@@
 The interaction term in the Hamiltonian becomes $\bm s_i \cdot \bm s_j$.
-A third spin model is the Heisenberg model, which is similar to the XY model except that $s_i$ is promoted to a three-dimensional unit vector.
-These models are all different in that they have progressively more spin degrees of freedom (DOFs), leading to progressively lower $T_c$.
+A third spin model is the Heisenberg model, which promotes $s_i$ to a three-dimensional unit vector.
+These models are all different in that they have progressively more spin DOFs, leading to progressively lower $T_c$.
 Each model also has specialized features, like the XY model's topologically protected vortices and the Heisenberg model's pseudo-phase transition \cite{tomita2014finite}, but these are beyond the scope of this report.
@@
 A spin model is most susceptible to magnetic field near $T_c$, because the presence of a magnetic field changes the value of $T_c$ and therefore can greatly change the magnetization of a system near $T_c$.
-In particular, the susceptibility diverges at $T_c$ and otherwise falls off to zero.
+Thus the susceptibility diverges at $T_c$ and otherwise falls off to zero.
 
@@
 The probability to go from $\phi$ to $\psi$ is the probability for the MCMC to be at state $\phi$ in the first place times the probability to switch to $\psi$.
-Thus, detailed balance requires
+Thus, reversibility requires
 \begin{equation}
@@
 \end{equation}
-Any method of generating a new state $\psi$ from the old state $\phi$ that satisfies Eq.~\ref{eqn:detailed-balance} is a valid MCMC.
 
@@
 At that point, the cluster's spins are flipped to form the new state $\psi$.
-Ref.~\cite{wolff1989collective} confirms that Eq.~\ref{eqn:wolff-algo} satisfies the detailed balance constraint and therefore is a good MCMC.
+Ref.~\cite{wolff1989collective} confirms that Eq.~\ref{eqn:wolff-algo} satisfies the reversibility constraint Eq.~\ref{eqn:detailed-balance} and therefore is a good MCMC.
 
@@
 
-We implemented the Wolff algorithm with a custom code base \footnote{\url{https://github.com/jack-dinsmore/quasing-model}} and executed for a variety of spin models and crystals.
+We implemented the Wolff algorithm with a custom code base\footnote{\url{https://github.com/jack-dinsmore/quasing-model}} and executed it for a variety of spin models and crystals.
 Except when stated otherwise, we used a finite-size lattice with initial conditions of random spins.
@@
   \includegraphics[width=0.49\linewidth]{../figs/penrose.pdf}
-  \caption{Magnetization and susceptibility as a function of $T_c$ under the XY and Ising models.
+  \caption{Magnetization and susceptibility as a function of $T_c$ under the Ising, XY, and Heisenberg models.
 \textit{Left}: A square lattice with $N=128^2=16,384$ sites.
@@
 The critical temperatures obtained from this approximation are shown in table \ref{tab:tc}.
-These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$ in units where $J=k_B=1$, and for the XY model MCMC methods give $T_c \approx 0.8816$.
-The Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which never have true phase transitions in any case.
+These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$ in units where $J=k_B=1$, and for the XY model larger lattice MCMC methods give $T_c \approx 0.8816$.
+The $N\rightarrow \infty$ Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which never have true phase transitions in any case.
 \begin{table}
@@
 The major difference between the two is the actual value of $T_c$, which is slightly higher for Penrose (table \ref{tab:tc}).
-This confirms the result of Ref.~\cite{penrose-ising}.
 
@@
 As expected, the result is symmetric around $\eta \rightarrow -\eta$.
-Anisotropy appears to increase $T_c$ near $\eta = 0$ in the Ising and XY models, though this trend turns over at some $\eta$ to achieve the $T_c\rightarrow 0$ value in the limit as $\eta \rightarrow \pm 1$.
+Anisotropy increases $T_c$ near $\eta = 0$ in the Ising and XY models, though this trend turns over at some $\eta$ to achieve $T_c\rightarrow 0$ in the limit as $\eta \rightarrow \pm 1$.
 For the Ising model, the turnover occurs at $\eta\approx \pm 0.625$ after which $T_c$ falls to zero rapidly.
@@
 
-If $\eta$ is shifted by some amount $\eta \rightarrow \eta - x$, then $T_c(\eta)$ possesses $\eta\rightarrow -\eta$ symmetry.
+If $\eta$ is shifted by some amount $\eta \rightarrow \eta - x$, then $T_c(\eta)$ possesses $\eta\rightarrow -\eta$ symmetry for the einstein tiling.
 This is surprising because $\eta$ and $-\eta$ correspond to completely different lattices, suggesting the presence of some greater symmetry in the einstein tiling which has not yet been remarked.
-Other models defined on the einstein tiling do not contain this symmetry.
+Other models defined on the einstein tiling do not always contain this symmetry.
 The shift value $x$ is 0.029 for the Ising model, 0.052 for the XY model, and 0.033 for the Heisenberg model, though these shift values may depend on both numerical fluctuations of the MCMC and on the finite $N$ used.
@@
 Quantum mechanical models such as the TIM cannot in general be represented by the same formalism as a classical model.
-However the the TIM is a special case: A $d$-dimensional TIM is equivalent to a $d+1$-dimensional Ising model, as first discovered in Ref.~\cite{schultz64two}.
+However the the TIM is a special case: A $d$-dimensional TIM is equivalent to a $d+1$-dimensional anisotropic Ising model, as first discovered in Ref.~\cite{schultz64two}.
 We will prove this in the case of $d=1$.
@@
 \begin{equation}
-  Z = \mathrm{Tr}\parens{e^{-\tau H}}.
+  Z = \mathrm{Tr}\parens{e^{-\beta H}}.
   \label{eqn:quantum-partition}
@@
 Explicit calculation verifies that $\braket{s_is_j|\bm \sigma_i \cdot \bm \sigma_j|s_i's_j'}=\delta_{s_i,s_j}\delta_{s'_i,s'_j}$.
-We sum over the $s'$ spins and replace the $s$ Kronecker delta with $s_i s_j$, which is equal to $\delta_{s_i,s_j}$ shifting by a constant.
+We sum over the $s'$ spins and replace the $s$ Kronecker delta with $s_i s_j$, which is equal to $\delta_{s_i,s_j}$ shifted by a constant.
 Likewise, $\braket{s|\sigma^3|s'}=\delta_{s,s'}\sim ss'$.
@@
 \begin{equation}
-  Z = 1 - \epsilon\sum_{\{s\},\{s'\}} \brackets{-J\sum_{\braket{ij}}\bm s_is_j - h\sum_is_i s'_i}.
+  Z = 1 - \epsilon\sum_{\{s\},\{s'\}} \brackets{-J\sum_{\braket{ij}} s_is_j - h\sum_is_i s'_i}.
 \end{equation}
@@
 Phase diagrams of the critical temperature as a function of the lattice anisotropy were shown; these are not predicted by Landau field theory and are therefore less theoretically constrained.
-Anisotropy was seen to increase the critical temperature of the square lattice, but decrease the critical temperature of the einstein tiling.
+Anisotropy was seen to increase the critical temperature of the square lattice but decrease the critical temperature of the einstein tiling.
 The rectangular lattice achieved maximum critical temperature at model-dependent ``turnover values'' of anisotropy, and the einstein tiling displayed unexpected symmetry after a model-dependent shift in the anisotropy.
